\chapter{Weakly étale subalgebras of fields}
\label{ch:mathlib-weakly-etale-subalgebra}

This chapter backs the new infrastructure file
\texttt{Proetale/Mathlib/RingTheory/WeaklyEtale/Subalgebra.lean}, born in
iter-021 as the Lane B pivot. It hosts the field-to-field specialisation of
the Stacks 092K subalgebra-descent lemma that the consumer
\texttt{Algebra.WeaklyEtale.isAlgebraic\_of\_isField\_aux} in
\texttt{FieldExtension.lean} requires to close the transcendental case of
Stacks 092P.

\section{Statement}

\begin{lemma}
  \label{lem:weakly-etale-adjoin-singleton}
  \lean{Algebra.WeaklyEtale.adjoin_singleton}
  \leanok
  Let $K$ be a field and $L$ a $K$-algebra with $K \to L$ weakly étale.
  For every $a \in L$, the $K$-subalgebra $K[a] = \mathrm{adjoin}_K\{a\}$
  is weakly étale over $K$.
  \uses{def:weakly-etale-algebra}
\end{lemma}

Recall that \emph{weakly étale} means $K$-flat together with formally
unramified. The flatness half is automatic: $K$ is a field, so every
$K$-module is free and hence flat, and in particular every $K$-subalgebra
of $L$ is $K$-flat. The substantive content is formal unramifiedness — or
equivalently formal étaleness, since the surjection
$\mu_{K[a]} : K[a] \otimes_K K[a] \to K[a]$ is a closed immersion of
finitely generated $K$-algebras.

\section{Strategy of proof}

\textit{Source: Stacks Project, tag 092K (specialised to field base).}

By \stacksproject{092K}, every $K$-subalgebra of a weakly étale
$K$-algebra is again weakly étale; the field-to-field case admits a
direct argument that avoids the general absolute-flatness descent
machinery. We outline the argument the formalisation should follow.

\begin{enumerate}
  \item \textbf{Reduce to formal étaleness of $\mu_{K[a]}$.} The
    multiplication $\mu_{K[a]} : K[a] \otimes_K K[a] \to K[a]$ is a
    surjective $K$-algebra homomorphism. By
    \texttt{Algebra.FormallyEtale.iff\_of\_surjective}
    (\stacksproject{02HN}), $\mu_{K[a]}$ is formally étale iff
    $\ker \mu_{K[a]}$ is an idempotent ideal of $K[a] \otimes_K K[a]$.
    Formally étale $\Rightarrow$ formally unramified, and flatness is
    automatic; together they give weakly étale.
  \item \textbf{Embed into $L \otimes_K L$.} Since $K$ is a field, the
    inclusion $K[a] \hookrightarrow L$ is faithfully flat as a map of
    $K$-modules (it is split injective). Hence the induced map
    \[
      K[a] \otimes_K K[a] \;\hookrightarrow\; L \otimes_K L
    \]
    is injective.
  \item \textbf{Inherit idempotency from $L \otimes_K L$.} The hypothesis
    $K \to L$ weakly étale gives that $\mu_L : L \otimes_K L \to L$ is
    flat with idempotent kernel
    (\Cref{lem:weakly-etale-preserves-absolutely-flat} +
    \Cref{lem:reduced-prime-maximal-of-absolutely-flat} machinery already
    formalised in
    \texttt{Proetale/Mathlib/RingTheory/WeaklyEtale/FieldExtension.lean}).
    Concretely, $\ker \mu_L = (e)$ for a single idempotent
    $e \in L \otimes_K L$.
  \item \textbf{The technical heart.} One shows that the inverse image of
    $(e)$ under the injection $K[a] \otimes_K K[a] \hookrightarrow
    L \otimes_K L$ is exactly $\ker \mu_{K[a]}$, and that this inverse
    image inherits idempotency. The Stacks argument routes through the
    explicit description of $\ker \mu_{K[a]}$ as the ideal generated by
    the diagonal elements $\{1 \otimes b - b \otimes 1 : b \in K[a]\}$.
    Each such generator already lies in $\ker \mu_L = (e)$, and the
    subring $K[a] \otimes_K K[a]$ is invariant under the multiplication
    by $e$ that exhibits idempotency in $L \otimes_K L$ (this last claim
    is the precise gap iter-020 documented: it requires more care than a
    naive idempotent lift).
  \item \textbf{Conclude.} Idempotency of $\ker \mu_{K[a]}$ in
    $K[a] \otimes_K K[a]$ promotes $\mu_{K[a]}$ to formally étale via
    \texttt{Algebra.FormallyEtale.iff\_of\_surjective}, and the field
    hypothesis on $K$ gives flatness, completing
    $\Algebra.\WeaklyEtale\ K\ K[a]$.
\end{enumerate}

The formalisation may begin from either signature (S2 above) or its
strengthened form (S1: every $K$-subalgebra of $L$ is weakly étale);
S2 is sufficient for the immediate downstream consumer and is
recommended as the entry point.

\section{Downstream consumer (informal)}

The bivariate contradiction inside
\texttt{isAlgebraic\_of\_isField\_aux} consumes
\Cref{lem:weakly-etale-adjoin-singleton} as follows. Suppose $a \in L$
is transcendental over $K$; then $K[a] \cong K[X]$ and
$K[a] \otimes_K K[a] \cong K[X, Y]$, which is a domain. By
\Cref{lem:weakly-etale-adjoin-singleton}, $K \to K[a]$ is weakly étale,
so $\mu_{K[a]}$ is formally étale and $\ker \mu_{K[a]}$ is principal,
generated by an idempotent of $K[X, Y]$. The only idempotents of a
domain are $0$ and $1$, so $\ker \mu_{K[a]} = (0)$ or $(1)$. But
$\ker \mu_{K[a]} = (X - Y)$, which is neither — contradiction. Hence
$a$ is algebraic.

This consumer wiring is deferred to iter-023; iter-022 closes the
descent step inside \Cref{lem:weakly-etale-adjoin-singleton}.

\section{iter-022 status}

iter-021 landed the file with two helper lemmas (FG of the kernel,
injectivity of the K-flat embedding $A\otimes_K A \hookrightarrow
L\otimes_K L$) and isolated the single substantive obstruction: the
descent of idempotency along that injection. iter-022's Lane B
closes that obstruction via the diagonal-generators argument
(Section "Strategy of proof" step (4) above), specifically:

\begin{enumerate}
  \item Extract the principal idempotent generator $e\in L\otimes_K L$
    of $\ker\mu_L$ using the project-local absolute-flatness
    infrastructure together with
    \texttt{Ideal.isIdempotentElem\_iff\_of\_fg}.
  \item Use $\mathrm{KaehlerDifferential.ideal\_eq\_diagonal}$ (or
    the equivalent in-Mathlib spelling) to expose
    $\ker\mu_A$ as the ideal generated by
    $\{1\otimes b - b\otimes 1 : b\in A\}$.
  \item Each diagonal of $A$ is also a diagonal of $L$, so its image
    under $i$ lies in $\ker\mu_L = (e)$. Hence $(1-e)$ annihilates
    each $i(1\otimes b - b\otimes 1)$.
  \item From $(1-e)\cdot \ker\mu_L = 0$ and the K-flat injection $i$,
    deduce $\ker\mu_A \cdot \ker\mu_A = \ker\mu_A$ inside
    $A\otimes_K A$. If a Mathlib "flat injective $\Rightarrow$
    idempotency descends" sub-lemma is unavailable, prove it locally
    as a private helper.
  \item Conclude
    $\Algebra.\WeaklyEtale\ K\ K[a]$ via
    \texttt{Ideal.Pure.of\_isIdempotentElem} $\to$
    \texttt{RingHom.Flat} of $\mu_A$.
\end{enumerate}
