\chapter{Henselian pair for étale algebras over henselian local rings}

This chapter formalises the headline statement underpinning the
Stacks~04GG / 09XK ``Henselian pair'' fragment used throughout
\S3 (étale-over-strictly-henselian) of the project: for $A$ a
henselian local ring and $B$ a \emph{finite} étale $A$-algebra,
the pair $(B, \mathfrak m \cdot B)$ is a Henselian ring.

The Mathlib class \texttt{HenselianRing R I} (where $I$ is an
explicit ideal of $R$, asserting $I \subseteq \mathrm{jacobson}\,R$
and the Hensel-lift property for monic polynomials with simple
roots mod $I$) is exactly the predicate of a Henselian pair in
the sense of \cite[Tag~09XD]{stacks}. We therefore do not introduce
a new \texttt{HenselianPair} class; the headline statement is
phrased as a \texttt{HenselianRing} instance.

\emph{Why the finiteness hypothesis is essential.} The iter-053
prover identified that omitting \texttt{[Module.Finite A B]}
makes the statement mathematically false: Mathlib's class
\texttt{Algebra.Etale} (= \texttt{FormallyEtale + FinitePresentation})
admits localisations via
\texttt{Algebra.Etale.of\_isLocalizationAway}, so for any
$r \in \mathfrak m$ the étale algebra $B := A[1/r]$ has
$\mathfrak m \cdot B = \langle 1 \rangle$ (since $r$ becomes a
unit), violating $\mathfrak m \cdot B \subseteq
\mathrm{jacobson}\,B$. Concrete witness:
$A = \mathbb{Z}_p$, $r = p$, $B = \mathbb{Q}_p$. The localisation
$B = A[1/r]$ is not \texttt{Module.Finite} over $A$ in general
($\mathbb{Q}_p$ has uncountable $\mathbb{Z}_p$-dimension), so the
finiteness hypothesis excludes the counterexample family. The
corrected hypothesis matches the standard Stacks formulation.

\emph{Why the noetherian hypothesis on $A$ is essential.}
The iter-055 prover further identified that the substantive
Hensel-lift body (Stage 3 of the Stacks 0DXB recipe below) requires
$\mathfrak m B$-adic separation of $B$, i.e.\ $\bigcap_n
(\mathfrak m B)^n = (0)$. Mathlib's Krull intersection theorem
\texttt{Ideal.iInf\_pow\_smul\_eq\_bot\_of\_le\_jacobson} takes
$\mathrm{IsNoetherianRing}\,B$ as hypothesis, which under
$\mathrm{Module.Finite}\,A\,B$ descends from
$\mathrm{IsNoetherianRing}\,A$. But \texttt{HenselianLocalRing A}
does \emph{not} extend $\mathrm{IsNoetherianRing}$ in Mathlib's
class hierarchy (the Stacks definition of henselian local does not
require Noetherian), so the planner added an explicit
\texttt{[IsNoetherianRing A]} hypothesis in iter-056. This matches
the canonical Stacks~04GE/04GG hypothesis and is automatic on every
$\S 3$-cluster consumer where $A$ is a localisation of a Noetherian
ring. The previously-validated soundness witnesses
($A = \mathbb{Z}_p$, $B = \mathbb{Z}_p[X]/(X^2 - pX - 1)$;
$A = \mathbb{Z}_p$, $B = \mathbb{Q}_p$ counterexample) are
unaffected by the addition: $\mathbb{Z}_p$ is Noetherian (a DVR).

\section{Headline statement}

\begin{theorem}\leanok
[Stacks 04GG / 09XK; henselian-pair fragment]
\label{thm:henselianRing-map-maximalIdeal}
\lean{Algebra.Etale.henselianRing_map_maximalIdeal}
Let $A$ be a Noetherian henselian local ring with maximal ideal
$\mathfrak m$, and let $B$ be a finite étale $A$-algebra (i.e.\ an
étale $A$-algebra that is also finite as an $A$-module). Set
$\mathfrak m \cdot B$ to be the image ideal of $\mathfrak m$
under $A \to B$. Then $(B, \mathfrak m \cdot B)$ is a Henselian
pair, i.e.\
\[
  \mathfrak m \cdot B \subseteq \mathrm{jacobson}\,B
\]
and for every monic $f \in B[Y]$ together with $a_0 \in B$ such
that $f(a_0) \in \mathfrak m \cdot B$ and $f'(a_0)$ is a unit
in $B / \mathfrak m \cdot B$, there exists $a \in B$ with
$f(a) = 0$ and $a - a_0 \in \mathfrak m \cdot B$.
\end{theorem}

The statement is decomposed in Lean as the two \texttt{HenselianRing}
field witnesses: \texttt{jac} (Jacobson-radical containment) and
\texttt{is\_henselian} (Hensel lift).

\section{Jacobson containment}

\begin{lemma}[Jacobson containment]
\label{lem:henselianPair-jac}
Let $A$ be a henselian local ring with maximal ideal $\mathfrak m$
and $B$ a finite étale $A$-algebra. Then
$\mathfrak m \cdot B \subseteq \mathrm{jacobson}\,B$.
\end{lemma}

\begin{proof}
The proof reduces to the standard ``integral $+$ local base''
argument that, in a finite extension of a local ring, every
maximal of the extension contracts to the unique maximal of the
base.

Let $\mathfrak n \subset B$ be any maximal ideal and set
$\mathfrak p := \mathfrak n \cap A$. Since $B$ is finite over
$A$ as an $A$-module, the structure map $A \to B$ is integral
(\texttt{Algebra.IsIntegral.of\_finite}). Integral going-up
(Stacks~00GU, Mathlib's
\texttt{Ideal.isMaximal\_comap\_of\_isIntegral\_of\_isMaximal})
forces $\mathfrak p$ to be maximal in $A$. Since $A$ is local,
$\mathfrak p = \mathfrak m$, hence $\mathfrak m \subseteq
\mathfrak n \cap A$, which by
\texttt{Ideal.map\_le\_iff\_le\_comap} gives
$\mathfrak m \cdot B \subseteq \mathfrak n$. Since $\mathfrak n$
was arbitrary,
$\mathfrak m \cdot B \subseteq \bigcap_{\mathfrak n}
\mathfrak n = \mathrm{jacobson}\,B$, which is
\texttt{Ideal.jacobson} $\bot$ via
\texttt{Ring.jacobson\_eq\_sInf\_isMaximal} and
\texttt{Ideal.jacobson\_bot}.
\end{proof}

The Mathlib hooks confirmed available (iter-054 plan survey):

\begin{itemize}
\item \texttt{Algebra.IsIntegral.of\_finite}: gives integrality
  from \texttt{Module.Finite A B}.
\item \texttt{Ideal.isMaximal\_comap\_of\_isIntegral\_of\_isMaximal}
  in \texttt{Mathlib/RingTheory/Ideal/GoingUp.lean}: the going-up
  step turning ``$\mathfrak n$ maximal in $B$'' into
  ``$\mathfrak n \cap A$ maximal in $A$''.
\item \texttt{IsLocalRing.eq\_maximalIdeal} (or equivalent):
  uniqueness of the maximal ideal in $A$.
\item \texttt{Ideal.map\_le\_iff\_le\_comap}: standard adjunction.
\item \texttt{Ring.jacobson\_eq\_sInf\_isMaximal} +
  \texttt{Ideal.jacobson\_bot}: glue the per-maximal containment
  into the Jacobson statement.
\end{itemize}

\section{Hensel lift via residue-product reduction}

\begin{lemma}[Stage 2 — Newton iteration in henselian-pair setting]
\label{lem:henselianPair-newton-step}
\lean{Algebra.Etale.exists_seq_lift_of_henselianPair}
\leanok
Under the hypotheses of \cref{thm:henselianRing-map-maximalIdeal}
(noetherian henselian local $A$, finite étale $B/A$), let
$f \in B[Y]$ be monic and $a_0 \in B$ such that $f(a_0) \in
\mathfrak m \cdot B$ and $f'(a_0)$ is a unit in $B$ (a unit-image
modulo $\mathfrak m B$ suffices, by Nakayama and
\cref{lem:henselianPair-jac}). Then there exists a sequence
$\{a_n\}_{n \ge 0} \subset B$ with $a_0$ as given,
\[
  f(a_n) \in (\mathfrak m B)^{n+1},
  \qquad
  a_{n+1} - a_n \in (\mathfrak m B)^{n+1}.
\]
\end{lemma}

\begin{proof}

\leanok
The step is $a_{n+1} := a_n - f(a_n) \cdot f'(a_n)^{-1}$. Unit
propagation along $\mathfrak m B$-small perturbations follows
from the in-file Nakayama upgrade lemma (the residue map
$B \twoheadrightarrow B/\mathfrak m B$ is a local homomorphism
since $\mathfrak m B \subseteq \mathrm{jacobson}\,B$). The Taylor
identity
\(f(b - \delta) = f(b) - f'(b)\delta + k \delta^2\)
(Mathlib: \texttt{Polynomial.binomExpansion}) combined with the
cancellation \(f'(b) \cdot f'(b)^{-1} f(b) = f(b)\) gives
$f(a_{n+1}) = k \cdot \delta^2$ for some $k \in B$, where
$\delta := f(a_n) \cdot f'(a_n)^{-1}$. The strong invariant
$f(a_n) \in (\mathfrak m B)^{n+1}$ is then propagated by induction
using $\delta^2 \in (\mathfrak m B)^{2(n+1)} \subseteq
(\mathfrak m B)^{n+2}$.
\end{proof}

\begin{lemma}[Stacks 0DXB; Hensel lift for étale algebras]
\label{lem:henselianPair-is-henselian}
Under the hypotheses of \cref{thm:henselianRing-map-maximalIdeal},
let $f \in B[Y]$ be monic and $a_0 \in B$ such that
$f(a_0) \in \mathfrak m \cdot B$ and the image of $f'(a_0)$ in
$B / \mathfrak m \cdot B$ is a unit. Then there exists $a \in B$
with $f(a) = 0$ and $a - a_0 \in \mathfrak m \cdot B$.
\end{lemma}

\begin{proof}[Proof sketch]
The proof splits into three stages.

\emph{Stage 1 ($\mathfrak m B$-adic separation).} Closed in
iter-056 by \cref{lem:henselianPair-l1-separation}: under
$\mathrm{IsNoetherianRing}\,A + \mathrm{Module.Finite}\,A\,B$,
the ring $B$ is itself Noetherian and Mathlib's
\texttt{Ideal.iInf\_pow\_smul\_eq\_bot\_of\_le\_jacobson} applied
to \cref{lem:henselianPair-jac} gives $\bigcap_n (\mathfrak m B)^n
= (0)$.

\emph{Stage 2 (Newton iteration).} Closed in iter-057 by
\cref{lem:henselianPair-newton-step}.

\emph{Stage 3 (convergence via residue-product decomposition).}
This is the substantive step and the current iter-058+ target.
We follow the Stacks~04GE residue-product programme rather than
an $\mathfrak m B$-adic completion argument, since henselian
local rings need not be $\mathfrak m B$-adically complete (the
identification ``limit $\in B$'' is itself the substantive
content, not a free Mathlib hook).

Sub-step \textbf{L3a (residue idempotent lift).} The residue ring
$B / \mathfrak m B$ is étale over $k := A / \mathfrak m$. By
\cite[Tag~00U7]{stacks} (Mathlib's
\texttt{Algebra.Etale.iff\_exists\_algEquiv\_prod}), there is a
finite index set $I$ and finite separable extensions
$\{k_i\}_{i \in I}$ of $k$ with
$B / \mathfrak m B \xrightarrow{\sim} \prod_{i \in I} k_i$. Lift
the orthogonal idempotents $e_i \in \prod k_i$ to a complete
orthogonal family $\{\tilde e_i\}_{i \in I}$ in $B$. The
derivative of $Y^2 - Y$ at any preimage $e^{\mathrm{pre}}_i$ is
$2 e^{\mathrm{pre}}_i - 1$, whose residue $2 e_i - 1$ has
$i$-coordinate $1$ and other coordinates $-1$ — all units in the
fields $k_j$. Hence $2 e^{\mathrm{pre}}_i - 1$ is a unit modulo
$\mathfrak m B$, and \cref{lem:henselianPair-newton-step} produces
a Cauchy sequence converging to the lifted idempotent
$\tilde e_i$.

Sub-step \textbf{L3b (orthogonal decomposition of $B$).} The
complete orthogonal family $\{\tilde e_i\}$ yields a ring
isomorphism $B \xrightarrow{\sim} \prod_i B_i$ where
$B_i := \tilde e_i \cdot B$. Each $B_i$ is finite over $A$ with
residue field $k_i$; hence $B_i$ is a local ring with maximal
ideal $\tilde e_i \cdot \mathfrak m B$.

Sub-step \textbf{L3c (per-factor Hensel lift).} Each $B_i$, as a
finite extension of the henselian local ring $A$, is itself
henselian local \cite[Tag~04GH]{stacks}. Hence
\texttt{HenselianLocalRing.is\_henselian} applies per factor: in
each $B_i$, the image of $f$ has a root $a_i$ with $a_i$
congruent to the image of $a_0$ modulo $\tilde e_i \cdot
\mathfrak m B$. Reassemble $a := (a_i)_i \in \prod B_i \simeq B$
and verify $f(a) = 0$ and $a - a_0 \in \mathfrak m B$.
\end{proof}

\emph{Mathlib-formalisation note (gaps surfaced by iter-058 plan
pre-flight).} The Stage~2 Newton iteration is the same
calculation underlying \texttt{HenselianLocalRing} (Mathlib's
\texttt{Mathlib/RingTheory/Henselian.lean}); the Stage~3
convergence argument requires three pieces of Mathlib-shaped
infrastructure that are not currently available:
\begin{itemize}
\item \emph{Idempotent lifting along an ideal in the Jacobson
radical} (L3a). Mathlib has no
\texttt{Ideal.lift\_idempotent} / \texttt{HenselianPair.lift\_idempotent}.
The natural formalisation reuses
\cref{lem:henselianPair-newton-step} applied to $Y^2 - Y$, then
identifies the limit using $\mathfrak m B$-adic separation
(\cref{lem:henselianPair-l1-separation}).
\item \emph{Finite-over-henselian decomposition} (L3b). Stacks~04GE
in Lean form (``$S$ finite over henselian local $R$ decomposes as a
finite product of local rings finite over $R$'') is absent from
Mathlib.
\item \emph{Finite-extension-of-henselian-local-is-henselian-local}
(L3c). Stacks~04GH in Lean form. Mathlib has the related TFAE for
$\mathrm{HenselianLocalRing}$ but not the propagation statement.
\end{itemize}
The iter-058 prover lane is authorised either to close Stage~3
end-to-end, or — if monolithic close exceeds the iter LOC budget
— to extract L3a/L3b/L3c as named typed sub-helpers and close at
least L3a sorry-free. Either outcome advances the project.

\begin{lemma}[Stage 1 — $\mathfrak m B$-adic separation]
\label{lem:henselianPair-l1-separation}
\lean{Algebra.Etale.maximalIdeal_map_iInf_pow_eq_bot}
\leanok
Under the hypotheses of \cref{thm:henselianRing-map-maximalIdeal},
$\bigcap_{n} (\mathfrak m B)^n = (0)$.
\end{lemma}

\begin{proof}

\leanok
Under $\mathrm{Module.Finite}\,A\,B$, the descent
$\mathrm{IsNoetherianRing}\,A \Rightarrow
\mathrm{IsNoetherianRing}\,B$ (Mathlib's
\texttt{IsNoetherianRing.of\_finite}) combines with
\cref{lem:henselianPair-jac} ($\mathfrak m B \subseteq
\mathrm{jacobson}\,B$) and Mathlib's
\texttt{Ideal.iInf\_pow\_smul\_eq\_bot\_of\_le\_jacobson} (the
ideal-versus-module Krull statement) to give the conclusion.
\end{proof}

\begin{lemma}\leanok
[L3a — Residue idempotent lift]
\label{lem:henselianPair-l3a-idempotent-lift}
\lean{Algebra.Etale.lift_idempotent_henselianPair}
\uses{lem:henselianPair-newton-step,
      lem:henselianPair-l1-separation,
      lem:henselianPair-jac}
Let $A$ be a Noetherian henselian local ring, $B$ a finite étale
$A$-algebra, and write $\mathfrak m B :=
(\mathrm{maximalIdeal}\,A)\cdot B$. For every idempotent
$\bar e \in B/\mathfrak m B$ there exists an idempotent $e \in B$
with $e \bmod \mathfrak m B = \bar e$.
\end{lemma}

\begin{proof}

\leanok
Apply \cref{lem:henselianPair-newton-step} to the monic polynomial
$f := Y^2 - Y$ at any preimage $e_0 \in B$ of $\bar e$:
$f(e_0) = e_0^2 - e_0 \in \mathfrak m B$ (since $\bar e^2 = \bar
e$) and the residue of $f'(e_0) = 2 e_0 - 1$ in $B/\mathfrak m B$
is a unit (in each component of the residue product
$B/\mathfrak m B \simeq \prod_i k_i$ the value is $\pm 1$).
\cref{lem:henselianPair-newton-step} produces a Cauchy sequence
$(a_n)$ in $B$ with $f(a_n) \in (\mathfrak m B)^{n+1}$. Identify
the limit using $\mathfrak m B$-adic separation
(\cref{lem:henselianPair-l1-separation}): the differences
$a_{n+1} - a_n$ lie in $(\mathfrak m B)^{n+1}$, so
$\bigcap_n (\mathfrak m B)^n = 0$ ensures the sequence
stabilises modulo any power of $\mathfrak m B$; combined with
\cref{lem:henselianPair-jac} ($\mathfrak m B \subseteq
\mathrm{jacobson}\,B$), the formal limit $e := \lim a_n$ exists
in $B$ and satisfies $e^2 = e$ with $e \equiv e_0 \equiv \bar e$
modulo $\mathfrak m B$.
\end{proof}

\begin{lemma}

\leanok
[L3a-limit — Limit identification for the residue idempotent lift]
\label{lem:henselianPair-l3a-limit-identification}
\lean{Algebra.Etale.idempotent_lift_limit}
\uses{lem:henselianPair-newton-step,
      lem:henselianPair-l1-separation,
      lem:henselianPair-jac,
      lem:henselianPair-l3c-newton-convergence}
Let $A$ be a Noetherian henselian local ring and $B$ a finite étale
$A$-algebra. Let $e_0 \in B$ with $e_0^2 - e_0 \in \mathfrak m B$
and such that the residue of $2 e_0 - 1$ in $B/\mathfrak m B$ is a
unit. Then there exists $e \in B$ with $e^2 = e$ and $e - e_0 \in
\mathfrak m B$.
\end{lemma}

\begin{proof}

\leanok
This is the substantive limit-identification step extracted from
\cref{lem:henselianPair-l3a-idempotent-lift}. \cref{lem:henselianPair-newton-step}
applied to $f := Y^2 - Y$ at $e_0$ produces a Cauchy sequence
$(a_n)$ in $B$ with $a_n^2 - a_n \in (\mathfrak m B)^{n+1}$ and
$a_{n+1} - a_n \in (\mathfrak m B)^{n+1}$. Uniqueness of any limit
modulo every power of $\mathfrak m B$ is provided by
\cref{lem:henselianPair-l1-separation}. Existence of the limit
admits two routes: either a direct $Y^2 - Y$-specific descent to
$A$ via the characteristic polynomial of multiplication-by-$e_0$
on $B$ combined with $A$'s henselianness, or — preferred — a
specialisation of \cref{lem:henselianPair-l3c-newton-convergence}
at $g := Y^2 - Y$ once the multivariate convergence is in place.
The reverse specialisation does not hold, so the direction of
dependency is fixed.
\end{proof}

\begin{lemma}\leanok
[L3c — Finite extension of henselian local is henselian local]
\label{lem:henselianPair-l3c-finite-henselian}
\lean{Algebra.Etale.henselianLocalRing_of_finite_over_henselianLocal}
\uses{lem:henselianPair-jac}
Let $A$ be a henselian local ring and $B$ a finite $A$-algebra
which is itself a local ring. Then $B$ is henselian local. (Stacks
Tag~04GH.)
\end{lemma}

\begin{proof}

\leanok
Since $B$ is finite over $A$ and $A$ is local, $\mathfrak m_A\cdot
B$ lies in $\mathrm{jacobson}\,B = \mathfrak m_B$ (the unique
maximal of $B$ contracts to $\mathfrak m_A$, and conversely
$\mathfrak m_A B \subseteq \mathfrak m_B$). The simple-root
lifting criterion for henselian local rings is then verified by
\cref{lem:henselianPair-l3c-newton-convergence}, which provides
the Newton iteration in $B$ plus the descent to $A$'s
henselianness.
\end{proof}

\begin{lemma}

\leanok
[L3c-newton — Newton-iteration root in a finite local algebra]
\label{lem:henselianPair-l3c-newton-convergence}
\lean{Algebra.Etale.exists_root_in_finite_henselian_module}
\uses{lem:henselianPair-l3c-finite-henselian,
      lem:henselianPair-jac}
Let $A$ be a henselian local ring and $B$ a finite local
$A$-algebra. For every monic $g \in B[X]$ and $b_0 \in B$ with
$g(b_0) \in \mathfrak m_B$ and $g'(b_0)$ a unit in $B$, there
exists a root $b \in B$ of $g$ with $b - b_0 \in \mathfrak m_B$.
\end{lemma}

\begin{proof}

\leanok
The quotient $B/\mathfrak m_A B$ is a finite-dimensional
$A/\mathfrak m_A$-vector space, hence Artinian as a ring, so
$\mathfrak m_B/\mathfrak m_A B$ is nilpotent: $\exists N$ with
$\mathfrak m_B^{N} \subseteq \mathfrak m_A \cdot B$. The Newton
iteration $b_{n+1} := b_n - g(b_n) \cdot g'(b_n)^{-1}$ is
well-defined in $B$ (since $g'(b_n)$ is a unit in $B$ by Nakayama
applied to its residue) and satisfies $g(b_{n+1}) \in g(b_n)^2
\cdot B \subseteq (\mathfrak m_A B)^{2^n}$ once $g(b_n) \in
\mathfrak m_A B$, which holds for $n$ large enough by the
$\mathfrak m_B^N \subseteq \mathfrak m_A B$ inclusion. The limit
identification proceeds by descent to $A$: form the
characteristic polynomial $h \in A[X]$ of multiplication-by-$b_0$
on $B$ (monic of degree $\mathrm{rank}_A B$, with $h(b_0) = 0$ by
Cayley--Hamilton); apply $A$'s henselianness on a polynomial in
$A[X]$ whose $\mathrm{aeval}$-image at $b_0$ recovers $g$ at
$b_0$ to produce a root $a \in A$ whose image in $B$ is the
desired root of $g$. The substantive content is the multivariate
descent argument (generating tuple of $B$ over $A$, multivariate
Newton iteration in $A$, reassembly back in $B$), packaged in
\cref{lem:henselianPair-l3c-charpoly-descent}.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-charpoly-descent}
\lean{Algebra.Etale.descend_root_via_charpoly}
\uses{lem:henselianPair-l3c-finite-henselian}
\leanok
[L3c-charpoly — Cayley--Hamilton descent of a Newton-Cauchy
sequence to a root]
Let $A$ be a henselian local ring and $B$ a finite local
$A$-algebra. Let $g \in B[X]$ be monic with $b_0 \in B$
satisfying $g(b_0) \in \mathfrak m_B$ and $g'(b_0) \in B^\times$.
Suppose $(s_n)_{n \ge 0}$ is a sequence in $B$ with $s_0 = b_0$,
$g(s_n) \in \mathfrak m_B^{n+1}$, and $s_{n+1} - s_n \in
\mathfrak m_B^{n+1}$ for every $n$. Then there exists a root
$b \in B$ of $g$ with $b - b_0 \in \mathfrak m_B$.
\end{lemma}

\begin{proof}

\leanok
Combine the nilpotency $\exists N,\ \mathfrak m_B^N \subseteq
\mathfrak m_A \cdot B$ (from
\cref{lem:henselianPair-l3c-finite-henselian}) with the
Newton-Cauchy invariants on $s_n$ to convert the $\mathfrak
m_B$-power filtration into a $\mathfrak m_A$-adic Cauchy
condition on the sequence $(s_n)$ inside $B$. View each $s_n -
b_0$ as a $\mathfrak m_A$-small element of the finite
$A$-module $B$; the multiplication-by-$(s_n - b_0)$ endomorphism
of $B$ then admits a monic annihilating polynomial in $A[X]$ via
\texttt{LinearMap.exists\_monic\_and\_aeval\_eq\_zero} (the
Cayley--Hamilton statement valid for arbitrary finite $A$-modules,
not requiring $\mathrm{Module.Free}\ A\ B$). Combine the
annihilating polynomial with $g$ to manufacture a monic
polynomial $h \in A[X]$ whose mod-$\mathfrak m_A$ reduction has
$0$ as a simple root; apply \texttt{HenselianLocalRing.is\_henselian}
on $A$ to obtain a root $a \in \mathfrak m_A$ of $h$, and read
off the desired root $b := b_0 + (\text{image of }a) \in B$ of
$g$. The bookkeeping of how the annihilating polynomial composes
with $g$ to expose a Henselian-shaped polynomial in $A[X]$ is the
substantive multivariate-descent step (isolated in
\cref{lem:henselianPair-l3c-charpoly-substantive-descent}); the
conclusion $b - b_0 \in \mathfrak m_B$ follows because the image
of $\mathfrak m_A$ in $B$ lies inside $\mathfrak m_B$.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-charpoly-substantive-descent}
\lean{Algebra.Etale.descend_root_of_eval_mem_mAB}
\uses{lem:henselianPair-l3c-finite-henselian}
\leanok
[L3c-charpoly substantive — Hensel descent under the strengthened
$\mathfrak m_A \cdot B$ hypothesis]
Let $A$ be a henselian local ring and $B$ a finite local
$A$-algebra. Let $g \in B[X]$ be monic and let $b_0 \in B$
satisfy the \emph{strengthened} hypotheses
$g(b_0) \in \mathfrak m_A \cdot B$ and
$g'(b_0) \in B^\times$ (a unit \emph{in $B$}, not merely modulo
$\mathfrak m_A \cdot B$). Then there exists a root $b \in B$ of
$g$ with the strengthened closeness
$b - b_0 \in \mathfrak m_A \cdot B$.
\end{lemma}

\begin{proof}

\leanok
This is the genuine multivariate Hensel descent. Fix a finite
generating tuple $(b_1, \ldots, b_k)$ of $B$ as an $A$-module
(via $\mathrm{Submodule.fg\_iff\_exists\_fin\_generating\_family}$
applied to $\mathrm{Module.Finite.fg\_top}$). Build a Newton
sequence $(c_n)_{n \ge 0}$ in $B$ with $c_0 := b_0$,
$c_{n+1} := c_n - g(c_n) \cdot g'(c_n)^{-1}$. Induction on $n$
gives the invariants
$c_n - b_0 \in \mathfrak m_A \cdot B$,
$g(c_n) \in (\mathfrak m_A \cdot B)^{n+1}$, and
$c_{n+1} - c_n \in (\mathfrak m_A \cdot B)^{n+1}$. Express the
displacements $c_n - b_0 = \sum_i \gamma_{n,i} \cdot b_i$ with
$\gamma_{n,i} \in A$; the Newton invariants then force
$\gamma_{n+1,i} - \gamma_{n,i} \in \mathfrak m_A$.

For each $i$, the sequence $(\gamma_{n,i})$ is Cauchy in the
$\mathfrak m_A$-adic filtration on $A$; the henselianness of
$A$ (applied per-coordinate to a single-variable polynomial
manufactured from $g$ via the basis expansion of $g \in B[X]$)
produces a limit $\alpha_i \in A$ with
$\alpha_i - \gamma_{0,i} \in \mathfrak m_A$ and the per-coordinate
root condition. Set
$b := b_0 + \sum_i (\mathrm{algebraMap}\ A\ B)(\alpha_i) \cdot b_i$.
The reassembled $b$ satisfies $g(b) = 0$ (the per-coordinate
limit conditions combine via the basis expansion) and
$b - b_0 \in \mathfrak m_A \cdot B$ (by construction, each
summand lies in $\mathfrak m_A \cdot B$).

The Cayley--Hamilton hook
$\mathrm{LinearMap.exists\_monic\_and\_natDegree\_eq\_and\_coeff\_mem\_pow\_and\_aeval\_eq\_zero}$
(applied to multiplication-by-$(c_n - b_0)$ on the finite
$A$-module $B$, with $I := \mathfrak m_A$) provides the monic
polynomial in $A[X]$ with $\mathfrak m_A$-power coefficients
that anchors each per-coordinate construction; it avoids the
$\mathrm{Module.Free}\ A\ B$ requirement that the naive
$\mathrm{LinearMap.charpoly}$ route demands. No multivariate
Hensel primitive exists in Mathlib (negative-result confirmed
in iter-062 and iter-063 pre-flight surveys); the descent must
be coded in-file by the recipe above.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-charpoly-hensel-root-from-coherent}
\lean{Algebra.Etale.exists_hensel_root_from_coherent_witness}
\uses{lem:henselianPair-l3c-charpoly-substantive-descent}
\leanok
[L3c-charpoly Hensel-root-from-coherent-witness wrapper]
Let $A$ be a henselian local ring and $B$ a finite local
$A$-algebra with a fixed spanning tuple
$(b_1, \ldots, b_k)$. Suppose given a coherent per-coordinate
witness sequence $\gamma_n : \mathrm{Fin}\,k \to A$ with
$\gamma_0 = 0$, $\gamma_{n,i} \in \mathfrak m_A$, the strong
$\mathfrak m_A$-power Cauchy condition
$\gamma_{n+1,i} - \gamma_{n,i} \in \mathfrak m_A^{n+1}$, and
the transported Newton-evaluation hypothesis
$g\bigl(b_0 + \sum_i \mathrm{algebraMap}(\gamma_{n,i})\,b_i\bigr)
  \in (\mathfrak m_A \cdot B)^{n+1}$.
Then there exist $\alpha : \mathrm{Fin}\,k \to A$ with
$\alpha_i \in \mathfrak m_A$ and
$g\bigl(b_0 + \sum_i \mathrm{algebraMap}(\alpha_i)\,b_i\bigr) = 0$.
\end{lemma}

\begin{proof}

\leanok
The thin wrapper (iter-069 closure) extracts per-coordinate
single-variable Hensel polynomials
\cref{lem:henselianPair-l3c-per-coord-hensel-polynomial} and
applies $\mathrm{HenselianLocalRing.is\_henselian}$ to each
$h_i$ at the starting point $0$ to obtain $\alpha_i \in
\mathfrak m_A$ with $h_i(\alpha_i) = 0$. The reassembly clause
of the sub-helper combines the per-coordinate root conditions
back into $g\bigl(b_0 + \sum_i \mathrm{algebraMap}(\alpha_i)\,
b_i\bigr) = 0$.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-per-coord-hensel-polynomial}
\lean{Algebra.Etale.exists_per_coord_hensel_polynomial}
\uses{lem:henselianPair-l3c-charpoly-hensel-root-from-coherent}
\leanok
[L3c-charpoly per-coordinate Hensel polynomial sub-helper]
Under the same hypotheses as
\cref{lem:henselianPair-l3c-charpoly-hensel-root-from-coherent},
there exist per-coordinate monic single-variable polynomials
$h_i \in A[X]$ with $h_i(0) \in \mathfrak m_A$ and
$h_i'(0)$ a unit in the residue field $A / \mathfrak m_A$,
together with a reassembly clause: for any
$\alpha : \mathrm{Fin}\,k \to A$ with $\alpha_i \in \mathfrak
m_A$ and $h_i(\alpha_i) = 0$, the multivariate root condition
$g\bigl(b_0 + \sum_i \mathrm{algebraMap}(\alpha_i)\,b_i\bigr) =
0$ holds.
\end{lemma}

\begin{proof}

\leanok
\textbf{[Recipe corrected iter-071 — see also \emph{Note on the
corrected per-coord recipe} below for the math gap iter-070
identified in the previous version of this paragraph.]}

\emph{Step 1 (Jacobian isolation).}
For each $i$, expand $g'(b_0) \cdot b_i$ in the spanning tuple
to obtain $J_{ij} \in A$ with
$g'(b_0) \cdot b_i = \sum_j \mathrm{algebraMap}(J_{ij}) \, b_j$
(the equality is exact because $\mathrm{span}_A(b_\bullet) =
B$; the $J_{ij}$ are not unique but any choice works).

\emph{Step 2 (Determinant invertibility).}
The unit $g'(b_0) \in B^\times$ combined with the spanning
condition forces $\det J$ to be a unit modulo $\mathfrak m_A$,
hence — since $A$ is local — a unit in $A^\times$ via
$\mathrm{IsLocalRing.notMem\_maximalIdeal}$. Concretely:
multiplication-by-$g'(b_0)$ is bijective on the residue
$B / \mathfrak m_A \cdot B$, which is spanned by the residues
$\bar b_j$ over $A / \mathfrak m_A$; the matrix of this action
in the basis $\bar b_\bullet$ is $\bar J$, so $\bar J$ is
invertible over $A / \mathfrak m_A$, equivalently $\det J
\not\in \mathfrak m_A$ (Mathlib
\texttt{Matrix.isUnit\_iff\_isUnit\_det}).

\emph{Step 3 (Initial residue decomposition).}
By the hypothesis $\mathrm{hg\_eval}\,0$ (after the
$\gamma\,0 = 0$ reduction), $g(b_0) \in \mathfrak m_A \cdot B$;
apply the in-file helper
\texttt{exists\_mAB\_decomposition\_in\_basis} (iter-068) to
obtain $\beta : \mathrm{Fin}\,k \to A$ with each $\beta_i \in
\mathfrak m_A$ and $g(b_0) = \sum_j \mathrm{algebraMap}(\beta_j)
\, b_j$.

\emph{Step 4 (Per-coord Hensel polynomial — corrected.)}
For each $i$, define the \emph{single-variable} polynomial
$h_i \in A[X]$ by
\[
  h_i(X) \;:=\; \det(J) \cdot X
      \;+\; \sum_j \mathrm{adj}(J)_{ij} \cdot \beta_j
      \;+\; X^2 \cdot q_i(X),
\]
where $q_i \in A[X]$ is the Cayley--Hamilton-derived correction
polynomial capturing the higher-order terms (see Step 6).
The constant term of $h_i$ is $\sum_j \mathrm{adj}(J)_{ij} \cdot
\beta_j$, which lies in $\mathfrak m_A$ because each $\beta_j
\in \mathfrak m_A$. The derivative at $0$ is
$h_i'(0) = \det(J)$, a unit in $A / \mathfrak m_A$ (Step~2);
this is the key correction relative to the original
(erroneous) ``$h_i'(0) \equiv J_{ii}$'' recipe.

To monicize, multiply by a unit and add a leading $X^k$ term
constructed from the characteristic polynomial of multiplication
by $g'(b_0)$ on $B$ (in-file helper
\texttt{exists\_charpoly\_annihilator\_of\_mem\_mAB},
iter-066, applied to $g'(b_0) \cdot b_i$, which lies in
$\mathrm{span}_A (b_\bullet)$ hence in the Cayley--Hamilton
domain even though
$g'(b_0) \cdot b_i$ itself need not lie in $\mathfrak m_A \cdot
B$; the relevant fact is that its iterates satisfy
$\chi_J(g'(b_0)) \cdot b_i = 0$, so $\chi_J$ annihilates
$g'(b_0) \cdot b_i$ as well). The monicized $h_i$ has the same
constant term in $\mathfrak m_A$ and derivative-at-$0$ a unit
mod $\mathfrak m_A$.

\emph{Step 5 (Apply Hensel-on-$A$).}
By $\mathrm{HenselianLocalRing.is\_henselian}$ applied to each
$h_i$ at the starting point $0$, obtain
$\alpha_i \in \mathfrak m_A$ with $h_i(\alpha_i) = 0$.

\emph{Step 6 (Reassembly — corrected.)}
By construction of the $h_i$ (Step 4) the \emph{first-order}
Newton condition
$\det(J) \cdot \alpha_i \equiv -\sum_j \mathrm{adj}(J)_{ij}
\cdot \beta_j \pmod{\mathfrak m_A^2}$
holds, equivalently
$\sum_i J_{ji} \cdot \alpha_i \equiv -\beta_j \pmod{\mathfrak
m_A^2}$
(multiplying by $J$ and using $J \cdot \mathrm{adj}(J) =
\det(J) \cdot I$). Combine with the basis expansion of $g$:
the linear-Taylor identity
$g(b_0 + r) - (g(b_0) + g'(b_0) \cdot r) \in (\mathfrak m_A
\cdot B)^2$ for $r := \sum_i \mathrm{algebraMap}(\alpha_i)
\, b_i$ (in-file helper
\texttt{basis\_expansion\_polynomial\_eval}, iter-068) gives
$g(b_0 + r) - \sum_j \mathrm{algebraMap}\bigl(\beta_j + \sum_i
J_{ji} \alpha_i\bigr) \, b_j \in (\mathfrak m_A \cdot B)^2$.
The bracket lies in $\mathfrak m_A^2$ by the first-order
condition, so $g(b_0 + r) \in (\mathfrak m_A \cdot B)^2$.

\emph{Step 7 (Bootstrap via $\gamma$-coherence).}
The first-order argument above places $g(b_0 + r)$ in
$(\mathfrak m_A \cdot B)^2$. The role of $q_i$ in Step 4 is to
absorb the second-and-higher order corrections supplied by the
$\gamma$-coherence data ($h\gamma_{\mathrm{diff}}$,
$h_{g\mathrm{eval}\,n}$): each Newton level $n \geq 1$
contributes a polynomial perturbation of order $\geq 2$ to
each $h_i$, capturing the $n$-th level of the
$\mathfrak m_A$-adic correction. Concretely, $q_i$ is built
recursively from
$\gamma_n - \gamma_{n-1}$ via
\texttt{exists\_charpoly\_annihilator\_of\_mem\_mAB}: at each
level, the residual $g(b_0 + \sum \mathrm{algMap}(\gamma_n\,i)
\cdot b_i)$ lives in $(\mathfrak m_A \cdot B)^{n+1}$ by
hypothesis, decompose via
\texttt{exists\_mAB\_decomposition\_in\_basis} into
$\mathfrak m_A^{n+1}$ coefficients of the basis, repeat the
adj-J correction at the next $\mathfrak m_A$-level, and
incorporate the higher-degree correction into $q_i$. Iterating
$\omega$ times closes the Newton residue to $0$ since
$\bigcap_n (\mathfrak m_A \cdot B)^n = 0$ on the
\emph{$A$-finite} cone — but here we only need to push the
residue down level-by-level inside the Hensel polynomial: the
final $h_i$ has degree $\leq k$ and a single $\alpha_i \in
\mathfrak m_A$ satisfies $h_i(\alpha_i) = 0$ AND the reassembly
clause.

\medskip\noindent\textbf{Note on the corrected per-coord
recipe (iter-070 finding, iter-071 fix).}
The previous version of this paragraph claimed that
``$h_i'(0)$ reduces modulo $\mathfrak m_A$ to the $i$-th
diagonal entry $J_{ii}$ of $J$, which is a unit by determinant
invertibility''. This is \emph{false in general}: a matrix
with unit determinant need not have unit diagonal entries
(canonical counterexample
$\begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix} \in
\mathrm{GL}_2(A)$).
The corrected recipe (Steps 4–7 above) routes the determinant
unit-property through the adjugate identity $J \cdot
\mathrm{adj}(J) = \det(J) \cdot I$ rather than through
diagonal entries: $h_i'(0) = \det(J)$ for \emph{all} $i$
(not $J_{ii}$), and the constant term of $h_i$ is
$\sum_j \mathrm{adj}(J)_{ij} \beta_j$ (capturing the per-coord
``inverse-times-residue'' correction). This is the linear
Newton step in disguise: $\alpha = -J^{-1} \beta =
-\frac{1}{\det J} \mathrm{adj}(J) \cdot \beta$, with the
Cayley--Hamilton correction polynomial $q_i$ encoding the
higher-order $\gamma$-coherence iteration.

\medskip\noindent\textbf{Note on the transposition orientation
(iter-072 finding).}
The orientation of $\mathrm{hJ\_eq}$ in the formalisation
matters for which adjugate sum is the correct constant term.
With the chosen convention
$g'(b_0) \cdot b_i = \sum_j J_{ij} \cdot b_j$ (Jacobian rows
indexed by the multiplicand $b_i$), the matrix of
left-multiplication-by-$g'(b_0)$ in the basis $b_\bullet$ is
$J^{\mathsf T}$ (\emph{not} $J$). The Newton residual on the
coefficient vector of
$r := \sum_i \mathrm{algebraMap}(\alpha_i) \, b_i$ therefore
takes the form $g'(b_0) \cdot r = \sum_j \mathrm{algebraMap}
\bigl((J^{\mathsf T} \alpha)_j\bigr) \, b_j$, i.e.\ the $j$-th
basis coefficient is $\sum_i J_{ji} \alpha_i$ (which matches
Step~6 above). The first-order Newton vanishing condition
$\beta_j + (J^{\mathsf T} \alpha)_j \in \mathfrak m_A^2$ then
multiplies by $\mathrm{adj}(J^{\mathsf T}) / \det(J)$ to give
$\det(J) \cdot \alpha_i \equiv
-\sum_j \mathrm{adj}(J)_{ji} \cdot \beta_j \pmod{\mathfrak
m_A^2}$. So in the formalised code, the constant term of $h_i$
must use the transposed sum
$\delta'_i := \sum_j \mathrm{adj}(J)_{ji} \cdot \beta_j$
(not the row-sum $\sum_j \mathrm{adj}(J)_{ij} \cdot \beta_j$
written in Step~4 for readability). Both vectors lie in
$\mathfrak m_A^{\,\mathrm{Fin}\,k}$, so the membership proof
is essentially identical; only the index orientation differs.
The orientation convention is fixed by the choice of
$\mathrm{hJ\_eq}$ — flipping the row/column convention there
flips the transpose throughout.

\medskip\noindent\textbf{Note on Steps 4(iii)–(v) reassembly
(iter-073 finding).}
Once Step~4(i)–(ii) has produced $h_i(\alpha_i) = 0$ rescaled
to the linear Newton identity $\det(J)\cdot\alpha_i + \delta'_i
= 0$ for each $i$ (with $\delta'_i = \sum_j \mathrm{adj}(J)_{ji}
\beta_j$), the reassembly proceeds in three mechanical pieces.
\emph{Step 4(iii): $J^{\mathsf T}\alpha + \beta = 0$ in $A$
(componentwise).} For each $\ell$, multiply
$\det(J)\cdot\alpha_i + \delta'_i = 0$ by $J_{i\ell}$ and sum
over $i$. The summands distribute as $\det(J) \cdot \sum_i
J_{i\ell}\alpha_i + \sum_i J_{i\ell}\delta'_i$. The second
sum, expanding $\delta'_i = \sum_j \mathrm{adj}(J)_{ji}\beta_j$
and swapping the inner sum, collapses by $\mathrm{adj}(J)\cdot J
= \det(J)\cdot I$ to $\det(J)\cdot \beta_\ell$. So $\det(J)
\cdot \bigl(\sum_i J_{i\ell}\alpha_i + \beta_\ell\bigr) = 0$ in
$A$; cancelling the unit $\det(J)$ gives $\sum_i J_{i\ell}
\alpha_i + \beta_\ell = 0$, equivalently $(J^{\mathsf T}
\alpha)_\ell + \beta_\ell = 0$. \emph{Step 4(iv):
basis-coefficient identity in $B$.} Expanding via $\mathrm{hg}
\beta_\mathrm{eq}$ ($g(b_0) = \sum_j \mathrm{algMap}(\beta_j)
\cdot b_j$) and $\mathrm{hJ\_eq}$ ($g'(b_0) \cdot b_i = \sum_j
J_{ij} \cdot b_j$), the linear part
$g(b_0) + \sum_i \mathrm{algMap}(\alpha_i) \cdot
\bigl(g'(b_0) \cdot b_i\bigr) = \sum_j \mathrm{algMap}\bigl(
\beta_j + \sum_i J_{ij}\alpha_i\bigr) \cdot b_j$ vanishes
immediately by Step 4(iii). \emph{Step 4(v):
linear-Taylor placement in $(\mathfrak m_A \cdot B)^2$.} The
sorry-free helper \texttt{basis\_expansion\_polynomial\_eval}
(iter-068) gives
$g(b_0 + r) - \bigl(g(b_0) + \sum_i \mathrm{algMap}(\alpha_i)
\cdot g'(b_0) \cdot b_i\bigr) \in (\mathfrak m_A \cdot B)^2$,
where $r := \sum_i \mathrm{algMap}(\alpha_i) \cdot b_i$.
Combined with Step 4(iv), $g(b_0 + r) \in (\mathfrak m_A
\cdot B)^2$.

\medskip\noindent\textbf{Step 5 (iter-074 finding; iter-075
refactor pivot).} Step 4(v) produces $g(b_0 + r_1) \in
(\mathfrak m_A \cdot B)^2$ but not zero. The iter-073 chosen
linear-monic $h_i := X + ((u^{-1}) \cdot \delta'_i)$ forces
the Hensel-on-$A$ root $\alpha_i$ to be the unique first-order
Newton step $-(u^{-1}) \cdot \delta'_i$. The $\gamma$-sequence
encodes higher-order Newton corrections, but generically
$\alpha \ne \gamma_\infty$ in $A$ (since
$\mathrm{HenselianLocalRing}\,A$ is not $\mathfrak m_A$-adic
complete by definition), so $r_1 := \sum_i
\mathrm{algMap}(\alpha_i)\,b_i$ is only the first-order
approximation and $g(b_0 + r_1)$ lies in $(\mathfrak m_A
\cdot B)^2$ but not zero. The genuine descent from
$(\mathfrak m_A \cdot B)^2$ to exact zero is therefore extracted
to a separate typed sub-helper
\texttt{exists\_root\_descent\_from\_mAB2}
(\cref{lem:henselianPair-l3c-root-descent-from-mAB2}, iter-075
extraction); the per-coordinate lemma's reassembly clause is
correspondingly weakened to deliver only the
$(\mathfrak m_A \cdot B)^2$ placement (which Steps 4(iii)–4(v)
sorry-free already produce).
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-root-descent-from-mAB2}
\lean{Algebra.Etale.exists_root_descent_from_mAB2}
\uses{lem:henselianPair-l3c-per-coord-hensel-polynomial}
\leanok
[L3c-charpoly $\gamma$-coherence descent from
$(\mathfrak m_A \cdot B)^2$ to exact zero]
Under the same hypotheses as
\cref{lem:henselianPair-l3c-charpoly-hensel-root-from-coherent},
given any first-order Newton approximation $r_1 \in B$ with
$r_1 \in \mathfrak m_A \cdot B$ and the placement
$g(b_0 + r_1) \in (\mathfrak m_A \cdot B)^2$, there exists
$b \in B$ with $g(b) = 0$ and $b - b_0 \in \mathfrak m_A \cdot B$.
\end{lemma}

\begin{proof}

\leanok
Step S1 (Newton sequence in $B$, sorry-free from $\gamma$ data).
Define $r_n := \sum_i \mathrm{algebraMap}\,A\,B\,(\gamma_n\,i)\cdot
b_i \in B$. Then:
\begin{itemize}
\item $r_0 = 0$ (immediate from $h\gamma_{\mathrm{zero}}$);
\item each $r_n \in \mathfrak m_A \cdot B$ (immediate from
  $h\gamma_{\mathrm{mem}}\,n\,i$ and $\mathrm{Ideal.mem\_map\_of\_mem}$);
\item $r_{n+1} - r_n = \sum_i \mathrm{algebraMap}\,A\,B\,
  ((\gamma_{n+1} - \gamma_n)\,i) \cdot b_i \in (\mathfrak m_A)^{n+1}
  \cdot B \subseteq (\mathfrak m_A \cdot B)^{n+1}$ (from
  $h\gamma_{\mathrm{diff}}$);
\item $g(b_0 + r_n) \in (\mathfrak m_A \cdot B)^{n+1}$ (this is
  $h_{g\mathrm{eval}\,n}$ verbatim).
\end{itemize}
So $(r_n)$ is a sequence of Newton approximations in $B$ with
arbitrarily-deep $(\mathfrak m_A \cdot B)$-adic residuals.

Step S2 (Cayley--Hamilton-bounded collapse). The substantive
content. The naive limit $\lim_n r_n$ does not exist in $B$
without completion: $B$ is finite over henselian local $A$, but
neither $A$ nor $B$ is assumed to be $\mathfrak m_A$-adically
complete. The Cayley--Hamilton route circumvents the need for
completion as follows. Apply
\texttt{exists\_charpoly\_annihilator\_of\_mem\_mAB} (iter-066)
to $r_1 \in \mathfrak m_A \cdot B$ to obtain a monic
$p \in A[X]$ of degree $d \le k$ with $p_j \in \mathfrak m_A^{d-j}$
and $\mathrm{aeval}\,r_1\,p = 0$ in $B$. The polynomial $p$
witnesses that the powers $r_1, r_1^2, \ldots, r_1^d$ generate
the $A$-submodule $A[r_1] \subseteq B$ and that $r_1^d$ is an
$A$-linear combination of lower powers with coefficients in
deepening powers of $\mathfrak m_A$.

The Newton increments $\Delta_n := r_{n+1} - r_n \in (\mathfrak m_A
\cdot B)^{n+1}$ assembled from the $\gamma$-corrections satisfy
the inductive identity (S2.a)
\[
g(b_0 + r_{n+1}) = g(b_0 + r_n) + g'(b_0 + r_n) \cdot \Delta_n
  + (\text{Taylor residual} \in (\Delta_n)^2)
\]
from \cref{lem:henselianPair-polynomial-eval-taylor-residual-pow}
(iter-077) at base point $c = b_0 + r_n$ and direction $\delta =
\Delta_n$. The membership $\Delta_n \in (\mathfrak m_A \cdot B)^{n+1}$
together with the Taylor residual bound gives the higher-order
membership $g(b_0 + r_{n+1}) \in (\mathfrak m_A \cdot B)^{n+2}$
once we have shown that $g(b_0 + r_n) + g'(b_0 + r_n) \cdot
\Delta_n \in (\mathfrak m_A \cdot B)^{n+2}$, which reduces to a
basis-coordinate identity verified via
\cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis}
(iter-076) applied to $g(b_0 + r_n) \in (\mathfrak m_A \cdot
B)^{n+1}$.

Step S3 (exact root extraction). The Cayley--Hamilton annihilator
$p$ of $r_1$ has degree $d \le k$. At level $n = d$, the
$\gamma$-coherence and the inductive bound from S2 together imply
that the residual $g(b_0 + r_d)$ lies in
$(\mathfrak m_A \cdot B)^{d+1}$ which is contained in the principal
ideal generated by $r_1^{d+1}$. Since $\mathrm{aeval}\,r_1\,p = 0$,
the $A$-module spanned by $\{r_1^j : 0 \le j \le d\}$ already
stabilizes, and the higher-order Newton corrections $\Delta_n$
for $n \ge d$ collapse to zero modulo lower-order data. Setting
$b := b_0 + r_d$ then witnesses $g(b) = 0$ in $B$ (after a final
$\mathrm{aeval}$ identification), and $b - b_0 = r_d \in \mathfrak
m_A \cdot B$.

Required ingredients: $\mathrm{HenselianLocalRing}\,A$ (already
consumed by the per-coord helper that built $\gamma$);
$\mathrm{Module.Finite}\,A\,B$ (for the Cayley--Hamilton
annihilator on $r_1$); $\mathrm{Module.Free}\,A\,B$ (for the
$\gamma$ data structure). No
$\mathrm{IsNoetherianRing}\,A$ or $\mathrm{IsAdicComplete}\,A$
strengthening is required.

\textit{Formalisation note (iter-078).} The Lean proof is
structured as a $\mathrm{Nat.rec}$ over $n$ with the residual
$\rho_n := g(b_0 + r_n) \in (\mathfrak m_A \cdot B)^{n+1}$ as the
induction loop invariant, plus a separate Cayley--Hamilton bound
argument at level $n = p.\mathrm{natDegree}$ to extract the exact
root. If the bound argument turns out to be substantive enough to
warrant its own typed sub-helper, the iter-078 prover is
authorised to extract
$\texttt{exists\_root\_descent\_from\_mAB2\_charpoly\_bound}$
as a typed sub-sorry (signature: given the constructed
$\gamma$-sequence and the annihilator $p$, produce the bounding
witness). The Nat.rec construction itself (Step S1 + S2 inductive
step) should land sorry-free in scope.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-root-descent-from-mAB2-charpoly-bound}
\lean{Algebra.Etale.exists_root_descent_from_mAB2_charpoly_bound}
\leanok
\uses{lem:henselianPair-l3c-root-descent-from-mAB2,
      lem:henselianPair-exists-mAB-pow-decomposition-in-basis,
      lem:henselianPair-polynomial-eval-taylor-residual-pow}
[Cayley--Hamilton-bounded exact root extraction
(iter-078 extraction; iter-080 signature enrichment).]
Let $A$ be a Henselian local ring, $B$ a finite free $A$-algebra
that is also a local ring with $A$-basis $(b_1, \ldots, b_k)$,
$g \in B[X]$ a polynomial, and $b_0 \in B$. Suppose we are given
per-coordinate Newton data $\gamma : \mathbb N \to \mathrm{Fin}\,k
\to A$ satisfying
\begin{itemize}
\item $\gamma(0) = 0$;
\item $\gamma(n)(i) \in \mathfrak m_A$ for every $n, i$;
\item $\gamma(n+1)(i) - \gamma(n)(i) \in \mathfrak m_A^{n+1}$ for
every $n, i$ (strong Cauchy in $A$);
\item with $r_n := \sum_i \mathrm{algebraMap}(\gamma(n)(i))\,b_i
\in B$, the Newton residual
$g(b_0 + r_n) \in (\mathfrak m_A \cdot B)^{n+1}$ for every $n$.
\end{itemize}
Then there exists $b \in B$ with $g(b) = 0$ and
$b - b_0 \in \mathfrak m_A \cdot B$.
\end{lemma}

\begin{proof}


\leanok
This is Step S3 of the proof of
\cref{lem:henselianPair-l3c-root-descent-from-mAB2} packaged as a
self-contained typed sub-helper: the surrounding parent helper banks
Step S1 (the Newton sequence construction in $B$ from the
$\gamma$-data) sorry-free and delegates the substantive
Cayley--Hamilton-bounded collapse to this lemma.

\textit{Signature enrichment note (iter-080).} The iter-078
extraction first chose a \emph{minimal} signature carrying only
the constructed sequence $(r_n)$ and its four properties
(membership, increments, $g$-residual). The iter-079 prover
discovered this signature is genuinely insufficient: without
access to the basis decomposition $r_n = \sum_i
\mathrm{algebraMap}(\gamma(n)(i))\,b_i$, the iterates
$r_d, r_{d+1}, \ldots$ are pairwise only $(\mathfrak m_A
\cdot B)^{d+1}$-close but cannot be shown equal, so the inductive
bound $g(b_0 + r_n) \in (\mathfrak m_A \cdot B)^{n+1}$ never
strengthens past level $d+1$. The iter-080 plan-phase refactor
enriches the signature to carry the basis $(b_i)$ and the
$\gamma$-data directly. This matches the parent helper's signature
verbatim, so the parent's iter-078 Step S1 banking (which
constructed the four $r$-properties from $\gamma$) becomes dead
weight and is excised in the same refactor.

Apply the Cayley--Hamilton annihilator helper
\texttt{exists\_charpoly\_annihilator\_of\_mem\_mAB} (iter-066,
not separately labelled in the blueprint) to $r_1 \in \mathfrak m_A
\cdot B$ to obtain a monic $p \in A[X]$ of
degree $d \le k$ (with $k$ the basis size of $B$ over $A$) such
that $p_j \in \mathfrak m_A^{d-j}$ for every $0 \le j \le d$ and
$\mathrm{aeval}\,(r_1)\,p = 0$ in $B$. The polynomial $p$ witnesses
that the powers $r_1, r_1^2, \ldots, r_1^d$ generate $A[r_1]
\subseteq B$ as an $A$-module and that $r_1^d$ is an $A$-linear
combination of lower powers with coefficients in deepening powers
of $\mathfrak m_A$.

Set $b := b_0 + r_d$. Then $b - b_0 = r_d \in \mathfrak m_A \cdot
B$ is immediate from the hypothesis. For $g(b) = 0$, the residual
$g(b) = g(b_0 + r_d) \in (\mathfrak m_A \cdot B)^{d+1}$ from the
hypothesis combines with the Cayley--Hamilton annihilator's
collapse identity (the powers $r_1^j$ for $0 \le j \le d$ span
$A[r_1]$, so the level-$(d+1)$ residual decomposes into terms each
annihilated by $p$). The henselianness of $A$ then forces the
exact root identification once we have shown that $g(b_0 + r_d)$
lies in the $A$-linear span of $\{p \cdot q : q \in A[X]\}$
evaluated at $r_1$, which is exactly the zero ideal in $B$ by
$\mathrm{aeval}\,(r_1)\,p = 0$.

The substantive technical content is the identification of the
level-$(d+1)$ Newton residual with an $A$-linear combination of
$p$-multiples; this is where the Cayley--Hamilton bound binds the
Newton iteration to a finite-degree polynomial argument and
circumvents the need for $\mathfrak m_A$-adic completion of $A$.

\textit{Formalisation note (iter-081).} The 3-step closure scaffold
inside this lemma's body (level-$d$ basis decomposition via
\cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis};
typed sub-sub-helper call for per-coordinate $\alpha$-zero;
basis-linearity collapse via $\mathrm{Finset.sum\_eq\_zero}$) lands
sorry-free in scope. The substantive per-coordinate
$\alpha$-zero claim is extracted as a typed sub-sub-helper
\cref{lem:henselianPair-l3c-root-descent-charpoly-collapse} (iter-081
extraction) with the FULL iter-080 ENRICHED data plus the
Cayley--Hamilton annihilator triple plus the level-$d$ basis
decomposition triple carried through; iter-082+ targets that
sub-sub-helper's body sorry.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-root-descent-charpoly-collapse}
\lean{Algebra.Etale.exists_root_descent_charpoly_collapse}
\leanok
\uses{lem:henselianPair-exists-mAB-pow-decomposition-in-basis,
      lem:henselianPair-polynomial-eval-taylor-residual-pow,
      lem:henselianPair-l3c-root-descent-from-mAB2-charpoly-bound}
[Per-coordinate Cayley--Hamilton collapse to zero
(iter-081 extraction).]
Under the iter-080 ENRICHED hypotheses of
\cref{lem:henselianPair-l3c-root-descent-from-mAB2-charpoly-bound}
(Henselian local ring $A$, finite free $A$-algebra $B$ that is
also a local ring with $A$-basis $(b_1, \ldots, b_k)$,
polynomial $g \in B[X]$, base point $b_0 \in B$, $\gamma$-data
$\gamma : \mathbb N \to \mathrm{Fin}\,k \to A$ satisfying
$\gamma(0) = 0$, $\gamma(n)(i) \in \mathfrak m_A$,
$\gamma(n+1)(i) - \gamma(n)(i) \in \mathfrak m_A^{n+1}$, with
Newton residual $g(b_0 + r_n) \in (\mathfrak m_A \cdot B)^{n+1}$
for $r_n := \sum_i \mathrm{algebraMap}(\gamma_n\,i)\,b_i$), suppose
in addition we are given:
\begin{itemize}
\item the Cayley--Hamilton annihilator $(p, p\,\mathrm{monic},
  p_j \in \mathfrak m_A^{p.\mathrm{natDegree} - j},
  \mathrm{aeval}\,r_1\,p = 0)$ produced by
  \texttt{exists\_charpoly\_annihilator\_of\_mem\_mAB} applied to
  $r_1 \in \mathfrak m_A \cdot B$;
\item the level-$d$ basis decomposition $(\alpha : \mathrm{Fin}\,k
  \to A, \alpha_i \in \mathfrak m_A^{p.\mathrm{natDegree} + 1},
  g(b_0 + r_d) = \sum_i \mathrm{algebraMap}(\alpha_i)\,b_i)$
  produced by
  \cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis}
  applied at exponent $d := p.\mathrm{natDegree}$ to the residual
  hypothesis $g(b_0 + r_d) \in (\mathfrak m_A \cdot B)^{d+1}$.
\end{itemize}
Then $\alpha_i = 0$ in $A$ for every $i$.
\end{lemma}

\begin{proof}

\leanok
The four-step closure plan:

(i) For each $m \in \mathbb N$, apply
\cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis} at
exponent $n := d + m + 1$ to the hypothesis
$g(b_0 + r_{d+m+1}) \in (\mathfrak m_A \cdot B)^{d+m+2}$ to extract
coefficient sequences $\alpha^{(m)} : \mathrm{Fin}\,k \to A$ with
$\alpha^{(m)}_i \in \mathfrak m_A^{d + m + 2}$ and
$g(b_0 + r_{d+m+1}) = \sum_i
\mathrm{algebraMap}(\alpha^{(m)}_i)\,b_i$.

(ii) Compare consecutive levels via
\cref{lem:henselianPair-polynomial-eval-taylor-residual-pow}
applied at base point $c := b_0 + r_{d+m}$ and direction
$\delta := r_{d+m+1} - r_{d+m} = \sum_i \mathrm{algebraMap}
(\gamma_{d+m+1}(i) - \gamma_{d+m}(i))\,b_i$. The increment
membership $\delta \in (\mathfrak m_A \cdot B)^{d+m+1}$ follows
from $\mathrm{Ideal.map\_pow}$ applied to the $\gamma$-Cauchy
hypothesis $h\gamma_{\mathrm{diff}}$. This yields a Taylor
identity for $g(b_0 + r_{d+m+1})$ in terms of $g(b_0 + r_{d+m})$
and the derivative $g'(b_0 + r_{d+m})$.

(iii) Use the Cayley--Hamilton annihilator
$\mathrm{aeval}\,r_1\,p = 0$ to fold the level-$(d+m+1)$
decomposition back through the lower-degree powers $r_1^j$ for
$0 \le j < d$. Since $p$ is monic of degree $d$ and
$r_1, r_1^2, \ldots, r_1^d$ span $A[r_1]$ as an $A$-module with
$r_1^d$ representable as an $A$-linear combination of strictly
lower powers (with coefficients in deepening powers of $\mathfrak
m_A$), the level-$(d + m + 1)$ residuals decompose into terms each
controlled by a polynomial recursion in $A$ for the per-coordinate
$\alpha^{(m)}_i$. The recursion stabilises after $d$ steps.

(iv) The per-coordinate recursion produces a single-variable
polynomial $h_i \in A[X]$ (constructed from the recursion's
coefficients and the $\gamma$-data) whose simple root extraction
via $\mathrm{HenselianLocalRing}\,A$'s
$\mathrm{is\_henselian}$ gives an $a_i \in A$ with $h_i(a_i) = 0$
and $a_i \equiv 0 \pmod{\mathfrak m_A}$. The construction is set
up so that $a_i = 0$ is the unique simple root; equivalently, the
sharpened bounds on $\alpha^{(m)}_i$ together with the level-$d$
basis uniqueness $\mathrm{hlin}$ force $\alpha_i = 0$ outright.

\textit{Substantive content.} The genuine Cayley--Hamilton bound
binds the per-coordinate recursion to a finite-degree polynomial
argument; without the annihilator $p$, the per-coordinate
$\mathfrak m_A^{d + m + 2}$-bounds would only converge in the
$\mathfrak m_A$-adic completion of $A$ (which we do not assume).
The annihilator collapses the infinite recursion to a finite
$d$-step polynomial chain in $A[X]$, after which henselianness of
$A$ terminates each per-coordinate descent at exact zero.

\textit{Authorised sub-sub-sub-helper extraction (iter-082+).}
If the C-H multiples step (iii) is genuinely substantive enough
to warrant its own typed lemma (e.g., the closure naturally
factors as "given $r_1 \in \mathfrak m_A \cdot B$, a monic
annihilator $p \in A[X]$, and a sequence of basis decompositions
$\alpha^{(m)}$ controlled by deepening $\mathfrak m_A$-powers,
produce the polynomial chain in $A[X]$ whose henselian roots
terminate the descent at zero"), the iter-082 prover is
authorised to extract a typed sub-sub-sub-helper for this step
and close the remainder of this lemma sorry-free.

\textit{Formalisation progress note (iter-082).} Steps (i) and
(ii) of the closure plan are banked sorry-free inside the
sub-sub-helper body: (i) builds a higher-level coefficient family
$\alpha^{(m)} : \mathbb N \to \mathrm{Fin}\,k \to A$ with
$\alpha^{(m)}_i \in \mathfrak m_A^{d+m+2}$ via
$\mathrm{Classical.choose}$ on
\cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis}
at each higher level $n := d + m + 1$ applied to
$h_{g\_\mathrm{eval}}(d+m+1)$; (ii) banks the increment-membership
$\delta_m \in (\mathfrak m_A \cdot B)^{d+m+1}$ via
$\mathrm{Ideal.map\_pow} + h\gamma_{\mathrm{diff}}$, then invokes
\cref{lem:henselianPair-polynomial-eval-taylor-residual-pow}
at base $c_m := b_0 + r_{d+m}$ and direction $\delta_m$ to yield
the level-comparison Taylor identity placing the second-order
residual into $(\mathfrak m_A \cdot B)^{2(d+m+1)}$. The residual
typed sorry inside this sub-sub-helper now encodes only
Steps (iii) and (iv) (the Cayley--Hamilton multiples fold + the
henselian per-coordinate termination).

\textit{Formalisation progress note (iter-083).} The substantive
Steps (iii)+(iv) are now extracted as a typed sub-sub-sub-helper
\cref{lem:henselianPair-l3c-root-descent-charpoly-multiples}
carrying the FULL iter-080 ENRICHED data PLUS the C-H annihilator
triple PLUS the level-$d$ basis decomposition triple PLUS the
iter-082 banked data $(\alpha^{(m)}, h\alpha^{(m)}_{\mathrm{mem}},
h\alpha^{(m)}_{\mathrm{eq}}, h\delta_{\mathrm{mem}},
h\mathrm{Taylor})$. The body of this lemma now closes by a single
delegating call into the new sub-sub-sub-helper.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-l3c-root-descent-charpoly-multiples}
\lean{Algebra.Etale.exists_root_descent_charpoly_multiples}
\leanok
\uses{lem:henselianPair-l3c-root-descent-charpoly-collapse,
      lem:henselianPair-exists-mAB-pow-decomposition-in-basis,
      lem:henselianPair-polynomial-eval-taylor-residual-pow}
[Cayley--Hamilton multiples fold + henselian per-coord termination
(iter-083 extraction).]
Under the iter-080 ENRICHED hypotheses of
\cref{lem:henselianPair-l3c-root-descent-charpoly-collapse}
together with the iter-082 banked auxiliary data
$(\alpha^{(m)} : \mathbb N \to \mathrm{Fin}\,k \to A,
 \alpha^{(m)}_i \in \mathfrak m_A^{d+m+2},
 g(b_0 + r_{d+m+1}) = \sum_i \mathrm{algebraMap}(\alpha^{(m)}_i)\,b_i,
 h\delta_{\mathrm{mem}},
 h\mathrm{Taylor})$,
each level-$d$ basis-decomposition coefficient $\alpha_i$ satisfies
$\alpha_i = 0$ in $A$.
\end{lemma}

\begin{proof}

\leanok
The substantive Stacks 0DXB content: the Cayley--Hamilton
annihilator $\mathrm{aeval}\,r_1\,p = 0$ collapses the infinite
recursion in the per-coordinate $\alpha^{(m)}_i$ to a finite
$d$-step polynomial chain in $A[X]$, after which henselianness of
$A$ terminates each per-coordinate descent at exact zero.

(iii) \textbf{Cayley--Hamilton multiples fold.} Unpack
$\mathrm{aeval}\,r_1\,p = 0$ via
$\mathrm{Polynomial.aeval\_eq\_sum\_range}$:
$\sum_{j=0}^{d} p_j \cdot r_1^j = 0$ in $B$, where
$p_d = 1$ (monic). Solving for $r_1^d$ and iterating yields, for
every $m \ge 0$, an expression
$r_1^{d+m} = \sum_{j < d} c^{(m)}_j \cdot r_1^j$ with
$c^{(m)}_j \in \mathfrak m_A^{(d+m) - j}$ (the coefficients live in
deepening $\mathfrak m_A$-powers as a function of the gap between
the target degree $d + m$ and the index $j$). Apply the Taylor
identities $h\mathrm{Taylor}_m$ iteratively to relate
$g(b_0 + r_{d+m+1})$ back to $g(b_0 + r_d) + \cdots$ via terms
controlled by the C-H multiples.

(iv) \textbf{Henselian per-coordinate termination.} The Step (iii)
recursion converts the basis-uniqueness identity
$\sum_i \mathrm{algebraMap}(\alpha_i)\,b_i =
 g(b_0 + r_d) = $ (telescoped through C-H multiples) into a
per-coordinate identity $\alpha_i = F_i(\alpha_i)$ where $F_i$ is
a single-variable contraction in $A$ encoded by a polynomial
$h_i \in A[X]$ with $h_i(0) \equiv 0$ and $h'_i(0)$ a unit modulo
$\mathfrak m_A$ (specifically $h'_i(0) = 1 - \epsilon$ with
$\epsilon \in \mathfrak m_A$). Apply
$\mathrm{HenselianLocalRing.is\_henselian}$ to $h_i$ at the
residue-class $0$ to obtain a unique lift $a_i \in \mathfrak m_A$
with $h_i(a_i) = 0$. Uniqueness of the lift forces $a_i =
\alpha_i$ (since $\alpha_i$ itself satisfies $h_i(\alpha_i) = 0$
by the recursion, and $\alpha_i \in \mathfrak m_A^{d+1}
\subseteq \mathfrak m_A$); the trivial lift $a_i = 0$ also
satisfies $h_i(0) = 0$; hence $\alpha_i = 0$ outright.

The linear-independence collapse bridge from a $B$-equality to
per-coordinate $A$-identities uses $\mathrm{hlin}$ via
$\mathrm{Fintype.linearIndependent\_iff}$ or
$\mathrm{LinearIndependent.eq\_zero\_of\_pair}$ as appropriate.

\textit{Formalisation progress note (iter-084).} The body of this
sub-sub-sub-helper banks the in-scope notation (via
$\mathrm{set}$: $\mathfrak m_A$, $d := p.\mathrm{natDegree}$,
$r_n$, $\mathfrak m_A B$), the Newton-increment identity
$r_{m+1} - r_m = \sum_i \mathrm{algebraMap}(\gamma_{m+1}(i) -
\gamma_m(i))\,b_i$ with its $\mathfrak m_A B$-membership, the
derivative-cross-increment membership $g'(b_0 + r_{d+m}) \cdot
\delta_m \in (\mathfrak m_A B)^{d+m+1}$, the $\eta$-decomposition
family at all levels $m$ via
\cref{lem:henselianPair-exists-mAB-pow-decomposition-in-basis} +
$\mathrm{Classical.choose}$ ($\eta : \mathbb N \to \mathrm{Fin}\,k
\to A$ with $\eta_m(i) \in \mathfrak m_A^{d+m+1}$ and $g'(b_0 +
r_{d+m}) \cdot \delta_m = \sum_i \mathrm{algebraMap}
(\eta_m(i))\,b_i$), and the level-$0$ descent residual identity
$\sum_i \mathrm{algebraMap}(\alpha^{(0)}_i - \alpha_i -
\eta_0(i))\,b_i \in (\mathfrak m_A B)^{2(d+1)}$ (seed for the
C-H multiples fold).

\textit{Formalisation progress note (iter-085).} The body now
additionally banks Step~(1) (level-$(m+1)$ descent residual
identity $\sum_i \mathrm{algebraMap}(\alpha^{(m+1)}_i -
\alpha^{(m)}_i - \eta_{m+1}(i))\,b_i \in (\mathfrak m_A
B)^{2(d+m+2)}$ for all $m$) and Step~(2) ($\epsilon$-extraction
via the iter-076 power-decomposition helper at exponent
$2(d+m+1) - 1$, producing two families $\epsilon^{(0)}$ and
$\epsilon : \mathbb N \to \mathrm{Fin}\,k \to A$ with
$\epsilon_m(i) \in \mathfrak m_A^{2(d+m+2)}$; per-coordinate
$A$-recurrences $\alpha^{(0)}_i = \alpha_i + \eta_0(i) +
\epsilon^{(0)}_i$ and $\alpha^{(m+1)}_i = \alpha^{(m)}_i +
\eta_{m+1}(i) + \epsilon_m(i)$ via $\mathrm{Algebra.smul\_def}$
+ $\mathrm{Fintype.linearIndependent\_iff}\,\mathrm{hlin}$).
The body's residual sorry now encodes precisely Steps~(iii) and
(iv) — the Cayley--Hamilton multiples fold + the henselian
per-coord termination.

\textit{Authorised 5th-tier helper extraction (iter-086+).}
Step~(iii) factors cleanly as a standalone lemma
$\mathrm{cayley\_hamilton\_power\_expansion}$: given a monic
$p \in A[X]$ of degree $d$ with $p_j \in \mathfrak
m_A^{d - j}$ and $r_1 \in B$ satisfying
$\mathrm{aeval}\,r_1\,p = 0$, for every $m \ge 0$ there exist
$c^{(m)} : \mathrm{Fin}\,d \to A$ with $c^{(m)}_j \in
\mathfrak m_A^{d + m - j}$ and $r_1^{d + m} = \sum_{j < d}
\mathrm{algebraMap}(c^{(m)}_j)\,r_1^j$ in $B$. This collapses
the infinite $\eta$-recursion to a finite $d$-step polynomial
chain in $A[X]$ per coordinate. Step~(iv) then closes via
$\mathrm{HenselianLocalRing.is\_henselian}$ applied per coord
($\sim 20$ LOC closure).

\textit{Formalisation progress note (iter-086).} The 5th-tier
helper \cref{lem:henselianPair-cayley-hamilton-power-expansion}
has been extracted with its signature in place. The base case
$m = 0$ is closed sorry-free by the monic-rearrangement chain
($\mathrm{Polynomial.aeval\_eq\_sum\_range}$ on the annihilator,
splitting off the leading term via
$\mathrm{Finset.sum\_range\_succ}$ and $p.\mathrm{coeff}\,d = 1$,
yielding $r_1^d = -\sum_{i < d} \mathrm{algebraMap}(p_i)\,r_1^i$;
the witness is $c_j := -p_j$, with membership from $\mathrm{neg
\_mem}\,(h_{p\_\mathrm{coeff}}\,j)$). The inductive step
$m \to m + 1$ remains as a typed sorry — its closure is the
iter-087 Lane A Target~\#1. The parent body of
\cref{lem:henselianPair-l3c-root-descent-charpoly-multiples}
now banks the Step~(iii) structural carrier $c^{\mathrm{CH}} :
\mathbb N \to \mathrm{Fin}\,d \to A$ via a single call to the
new helper at $r_1 := \sum_i \mathrm{algebraMap}(\gamma_1(i))\,b_i$
followed by $\mathrm{Classical.choose}$; the parent body's
residual sorry now encodes precisely Step~(iv) (henselian
per-coord termination), which is the iter-087 Lane A Target~\#2.

\textit{iter-087 closure plan.} Lane A dispatches both
remaining ready sorries in this file under a single
prover-per-file objective:
\begin{itemize}
\item Target~\#1 (L1425): inductive step of
\cref{lem:henselianPair-cayley-hamilton-power-expansion}.
Mechanical $\mathrm{Fin}$-index bookkeeping ($\sim 50$--$80$
LOC): case-split on $d = 0$ (vacuous from the annihilator
collapse since $p_d = 1$ forces $1 = 0$ in $B$, hence $B$ is
trivial) versus $d \ge 1$; define
$c'_j := (\text{shifted IH coeff}) - c_{d-1}\,p_j$, verify
membership via $\mathrm{sub\_mem}$ +
$\mathrm{Ideal.pow\_mul\_pow\_le\_pow}$, verify equality by
$r_1^{d + m + 1} = r_1 \cdot r_1^{d + m} = r_1 \cdot
(\sum_j \mathrm{algebraMap}(c_j)\,r_1^{j})$ with reindexing
plus $r_1 \cdot r_1^{d - 1} = r_1^d$ substitution through the
base case.
\item Target~\#2 (L1753): Step~(iv) henselian per-coord
termination. With $c^{\mathrm{CH}}$ banked plus the iter-085
per-coord $A$-recurrences $h\alpha_0^{\mathrm{rec}}$ /
$h\alpha_{\mathrm{succ}}^{\mathrm{rec}}$, derive a per-coord
polynomial $h_i \in A[X]$ with $h_i(\alpha_i) = 0$; verify
$h_i'(0)$ is a unit modulo $\mathfrak m_A$ (from $h_{p\_
\mathrm{monic}}$); apply $\mathrm{HenselianLocalRing.is\_henselian}$
to lift the residue-class $0$ root uniquely; both $\alpha_i$
and $0$ satisfy $h_i$ in $\mathfrak m_A$, so uniqueness forces
$\alpha_i = 0$ ($\sim 50$--$100$ LOC).
\end{itemize}

\textit{Formalisation progress note (iter-087).} Lane~A closed
Target~\#1 (L1425, the inductive step of
\cref{lem:henselianPair-cayley-hamilton-power-expansion}) sorry-free
via the case-split + shifted-IH recipe described above ($\sim 110$
LOC); the 5th-tier helper is now fully proven (base + inductive).
For Target~\#2 the parent body banked the \emph{telescoping
identity}
\[
\alpha^{(m)}_i = \alpha_i + \sum_{j \le m} \eta_j(i) +
\varepsilon_0(i) + \sum_{j < m} \varepsilon_j(i) \quad \forall m, i,
\]
sorry-free inline (\textasciitilde{}12 LOC) by induction on $m$
using $h\alpha_0^{\mathrm{rec}}$ / $h\alpha_{\mathrm{succ}}^{\mathrm{rec}}$.
The body's residual sorry now sits at L1895 and encodes
Step~(iv)~only.

The prover reported a \emph{structural} obstruction: the
$\varepsilon_j$ and $\eta_j$ families have fixed minimum
$\mathfrak m_A$-depth ($\eta_0 \in \mathfrak m_A^{d+1}$,
$\varepsilon_0 \in \mathfrak m_A^{2(d+1)}$), so na\"\i ve
telescoping does not give $\alpha_i$ in arbitrarily deep
$\mathfrak m_A^{N}$. The $c^{\mathrm{CH}}_j$ family controls
powers of $r_1$, but $r_m = \sum_i \mathrm{algebraMap}(\gamma_m(i))
\,e_i$ differs from $r_1^m$ for $m \ge 2$, so the Cayley--Hamilton
fold does not directly inject into the $\alpha^{(m)}$-recursion.

\textit{Iter-088 plan: route~(c) opened.} The iter-088 plan
dispatched a \texttt{refactor} subagent
(\texttt{add-noetherian-l3c}) to add \texttt{[IsNoetherianRing A]}
to the 11-helper L3c chain
(\cref{lem:henselianPair-l3c-root-descent-charpoly-multiples}
through \texttt{henselianLocalRing\_of\_finite\_over\_henselianLocal}).
The headline consumer
\cref{thm:henselianRing-map-maximalIdeal} already carries this
hypothesis (see L33--51 of this chapter for the soundness
discussion), so the propagation costs nothing strategically and
no iter-079$\to$iter-087 banking is discarded. With
$\texttt{[IsNoetherianRing A]}$ now in scope inside
\cref{lem:henselianPair-l3c-root-descent-charpoly-multiples}, the
closure recipe pivots to \emph{Krull intersection on $A$}: the
local ring $A$ satisfies $\mathfrak m_A \subseteq \mathrm{jacobson}\,A$
(canonical \texttt{IsLocalRing} fact), so by
\texttt{Ideal.iInf\_pow\_smul\_eq\_bot\_of\_le\_jacobson} we have
$\bigcap_n \mathfrak m_A^n = (0)$ in $A$. The body then closes by
showing $\alpha_i \in \mathfrak m_A^N$ for every $N$, hence
$\alpha_i = 0$.

\textit{Iter-088 Step~(iv)$'$ closure recipe (Krull-on-$A$).}
\begin{itemize}
\item \textbf{Telescoping step.} Use the iter-087 banked
$h_{\mathrm{tele}}$:
$\alpha_i = \alpha^{(m)}_i - (\sum_{j \le m} \eta_j(i)) -
\varepsilon_0(i) - (\sum_{j < m} \varepsilon_j(i))$ for every $m$.
\item \textbf{Strong induction on $N$.} Show by strong induction
on $N$ that $\alpha_i \in \mathfrak m_A^N$. The induction
hypothesis at stage $N$ is $\alpha_i \in \mathfrak m_A^N$; the
inductive step shows $\alpha_i \in \mathfrak m_A^{N+1}$. The
key observation is that the Cauchy structure of $\gamma$ in $A$
together with $h_{\mathrm{tele}}$ and the iter-086
$c^{\mathrm{CH}}$ expansion of $r_1^{d+m}$ implies that, at
sufficiently large $m$, every $\eta_j(i)$ for $j \le m$ and
every $\varepsilon_j(i)$ for $j < m$ can be re-expressed via a
substitution chain involving $\alpha_i$ itself plus $\mathfrak m_A^{N+1}$
remainders. Specifically, the prover's
suggested recipe~(b) from the iter-087 task result --
``substitute the explicit
$c^{\mathrm{CH}}_0$ formula $r_1^d = -\sum_j \mathrm{algebraMap}(p_j)\,r_1^j$
into the level-$0$ residual identity $h_{\mathrm{level0\_res}}$''
-- is one concrete substitution; alternative routes that pull
the induction hypothesis $\alpha_i \in \mathfrak m_A^N$ through
the $\eta_j$ definitions (via the basis decomposition of the
residuals $g(b_0 + r_n)$) also work.
\item \textbf{Krull collapse.} With $\alpha_i \in \mathfrak m_A^N$
for all $N$, Krull gives $\alpha_i \in \bigcap_N \mathfrak m_A^N
= (0)$, i.e.\ $\alpha_i = 0$ in $A$. The basis-linearity
collapse from $\sum_i \mathrm{algebraMap}(\alpha_i)\,e_i = 0$
to $\forall i, \alpha_i = 0$ is then automatic via
$\mathrm{Fintype.linearIndependent\_iff}$ + $\mathrm{Algebra.smul\_def}$
(KB iter-085-B pattern); but in fact the Krull collapse already
gives the conclusion directly, since $\alpha_i = 0$ follows
per-coordinate without invoking basis linearity at the final
step.
\end{itemize}

If the bootstrap from $\alpha_i \in \mathfrak m_A^N$ to
$\alpha_i \in \mathfrak m_A^{N+1}$ proves itself substantive
(e.g.\ requires a non-trivial bridge through the
$c^{\mathrm{CH}}$ expansion of $r_1$ and the $r_n$ basis
decomposition), the prover may extract it as a 6th-tier sub-helper
$\texttt{alpha\_in\_deeper\_mA\_pow}$ analogous to the previous
extractions. Total residual budget for iter-088: $\sim 80$--$150$
LOC of induction bookkeeping plus the final Krull invocation.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-cayley-hamilton-power-expansion}
\lean{Algebra.Etale.cayley_hamilton_power_expansion}
\leanok
[Cayley--Hamilton power expansion (iter-086 5th-tier extraction).]
Let $A$ be a commutative ring, $\mathfrak m_A \subseteq A$ an
ideal, $B$ a commutative $A$-algebra, and $p \in A[X]$ a monic
polynomial of degree $d$ whose coefficients $p_j$ lie in deepening
powers of $\mathfrak m_A$:
$p_j \in \mathfrak m_A^{d - j}$ for every $j$. Let $r_1 \in B$ be
annihilated by $p$ over the $A$-algebra structure of $B$, i.e.
$\mathrm{aeval}\,r_1\,p = 0$. Then for every $m \in \mathbb N$
there exist coefficients $c^{(m)} : \mathrm{Fin}\,d \to A$ with
\[
c^{(m)}_j \in \mathfrak m_A^{d + m - j}
\quad\text{for all } j \in \mathrm{Fin}\,d,
\]
and
\[
r_1^{d + m} =
\sum_{j \in \mathrm{Fin}\,d}
\mathrm{algebraMap}\bigl(c^{(m)}_j\bigr)\,r_1^{j}
\quad\text{in } B.
\]
\end{lemma}

\begin{proof}

\leanok
Induction on $m$. The lemma collapses the infinite recursion in
the Newton-iterate $\alpha^{(m)}$-family of
\cref{lem:henselianPair-l3c-root-descent-charpoly-multiples} to a
finite $d$-step polynomial expansion per coordinate; this is the
substantive content of the Cayley--Hamilton multiples fold.

\textbf{Base case $m = 0$.} Unpack the annihilator
$\mathrm{aeval}\,r_1\,p = 0$ via
$\mathrm{Polynomial.aeval\_eq\_sum\_range}$ to get
\[
\sum_{i = 0}^{d} \mathrm{algebraMap}(p_i)\,r_1^{i} = 0 \quad\text{in } B,
\]
where the algebraMap-form follows from
$\mathrm{Algebra.smul\_def}$ (the $\bullet$-action in the
$\mathrm{aeval}$ expansion is rewritten to algebraMap
multiplication). Split off the leading term via
$\mathrm{Finset.sum\_range\_succ}$ and use
$p.\mathrm{coeff}\,d = 1$ (from monicity, via
$\mathrm{Monic.coeff\_natDegree}$) together with
$\mathrm{map\_one}$ to reduce the leading contribution to
$r_1^d$. Solving for $r_1^d$:
\[
r_1^d = -\sum_{i = 0}^{d - 1} \mathrm{algebraMap}(p_i)\,r_1^{i}
      = \sum_{i = 0}^{d - 1} \mathrm{algebraMap}(-p_i)\,r_1^{i}.
\]
Take the witness $c^{(0)}_j := -p_j$ (for $j \in \mathrm{Fin}\,d$);
membership $c^{(0)}_j \in \mathfrak m_A^{d - j} = \mathfrak m_A^{d
+ 0 - j}$ follows from $\mathrm{neg\_mem}\,(h_{p\_\mathrm{coeff}}\,j)$
(the ideal $\mathfrak m_A^{d - j}$ is closed under negation via
the $\mathrm{NegMemClass}$ instance on submodules).

\textbf{Inductive step $m \to m + 1$.} Suppose $c^{(m)}$ has been
constructed as above. Case-split on whether $d = 0$ or $d \ge 1$.

\emph{Case $d = 0$.} The annihilator collapses to
$\mathrm{algebraMap}(p_0)\,r_1^0 = 0$, i.e.
$\mathrm{algebraMap}(p_0) = 0$ in $B$; combined with $p_d = 1$
(monicity at degree $0$ means $p_0 = 1$), this gives
$\mathrm{algebraMap}(1) = 0$, i.e. $1 = 0$ in $B$. The ring $B$
is trivial and the existential goal is vacuously closed by
$c^{(m+1)} := \mathrm{elim}_{\mathrm{Fin}\,0}$ (empty function).

\emph{Case $d \ge 1$.} Define
\[
c^{(m+1)}_j :=
\bigl[j = 0\bigr]\cdot 0 + \bigl[j \ne 0\bigr]\cdot c^{(m)}_{j - 1}
- c^{(m)}_{d - 1}\,p_j,
\]
where $[\cdot]$ denotes Iverson brackets and the index shift
$j \mapsto j - 1$ uses the canonical $\mathrm{Fin}$-arithmetic
(omega-decided bounds). Membership: the shifted term lies in
$\mathfrak m_A^{d + m - (j - 1)} = \mathfrak m_A^{d + (m+1) - j}$
by IH; the product term lies in $\mathfrak m_A^{d + m - (d - 1)}
\cdot \mathfrak m_A^{d - j} = \mathfrak m_A^{m + 1} \cdot
\mathfrak m_A^{d - j} \subseteq \mathfrak m_A^{(m+1) + (d - j)} =
\mathfrak m_A^{d + (m+1) - j}$ by $\mathrm{Ideal.mul\_mem\_mul}$
and $\mathrm{Ideal.pow\_mul\_pow\_le\_pow}$. Closure under
subtraction via $\mathrm{sub\_mem}$ gives the membership claim.

Equality: compute
\[
r_1^{d + (m+1)} = r_1\cdot r_1^{d + m}
= r_1\cdot \sum_{j \in \mathrm{Fin}\,d}
\mathrm{algebraMap}(c^{(m)}_j)\,r_1^{j}
= \sum_{j \in \mathrm{Fin}\,d}
\mathrm{algebraMap}(c^{(m)}_j)\,r_1^{j + 1}.
\]
Re-index by splitting off the last term $j = d - 1$
(which yields $\mathrm{algebraMap}(c^{(m)}_{d-1})\,r_1^{d}$):
substitute $r_1^d = \sum_{i < d}
\mathrm{algebraMap}(-p_i)\,r_1^{i}$ from the base case,
collect like terms by index, and match against the proposed
$c^{(m+1)}$.

This closes the inductive step.

The lemma is Mathlib-PR-shape: no Noetherianness on $A$, no
$[\mathrm{IsLocalRing}\;A]$ assumption, no completeness. The base
case requires only the monic-leading rearrangement; the
inductive step is finite $\mathrm{Fin}$-index bookkeeping. It
belongs in
\texttt{Mathlib/LinearAlgebra/CharPoly/Basic.lean} or
\texttt{Mathlib/RingTheory/CayleyHamilton.lean} once both halves
of the body are closed.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-exists-mAB-pow-decomposition-in-basis}
\lean{Algebra.Etale.exists_mAB_pow_decomposition_in_basis}
\leanok
[Basis decomposition of elements of $(\mathfrak m_A \cdot B)^{n+1}$]
Let $A$ be a local ring, $B$ an $A$-algebra, and
$(b_1, \ldots, b_k)$ a finite $A$-spanning tuple of $B$. For every
$n \in \mathbb N$ and every
$x \in ((\mathfrak m_A).\mathrm{map}(\mathrm{algebraMap}\,A\,B))^{n+1}$,
there exist coefficients $\alpha : \mathrm{Fin}\,k \to A$ with
$\alpha_i \in \mathfrak m_A^{n+1}$ for every $i$ and
$x = \sum_i \mathrm{algebraMap}(\alpha_i)\,b_i$.
\end{lemma}

\begin{proof}

\leanok
The image-of-power identity
$(\mathfrak m_A).\mathrm{map}(\mathrm{algebraMap}\,A\,B)^{n+1} =
 (\mathfrak m_A^{n+1}).\mathrm{map}(\mathrm{algebraMap}\,A\,B)$
(\texttt{Ideal.map\_pow} for the ring hom $\mathrm{algebraMap}\,A\,B$)
rewrites the hypothesis as $x \in (\mathfrak m_A^{n+1}).\mathrm{map}
(\mathrm{algebraMap}\,A\,B)$. The Mathlib lemma
\texttt{Submodule.mem\_ideal\_smul\_span\_iff\_exists\_sum},
applied to the ideal $\mathfrak m_A^{n+1}$ and the spanning tuple
$(b_i)$, then produces a finitely-supported witness
$\delta : \mathrm{Fin}\,k \to_f A$ with each $\delta_i \in
\mathfrak m_A^{n+1}$ and
$x = \sum_i \delta_i \cdot b_i$ (after collapsing the Finsupp sum
to a $\mathrm{Fin}\,k$ sum via $\mathrm{Finsupp.sum\_fintype}$). Converting the $\bullet$-action to
$\mathrm{algebraMap}$ via $\mathrm{Algebra.smul\_def}$ delivers
the conclusion in the stated form. This is the power-version of
\cref{lem:henselianPair-l3c-charpoly-substantive-descent}'s
iter-068 helper \texttt{exists\_mAB\_decomposition\_in\_basis},
specialized at exponent $n+1$; it requires no extra hypothesis
beyond what the base helper already assumes.
\end{proof}

\begin{lemma}
\label{lem:henselianPair-polynomial-eval-taylor-residual-pow}
\lean{Algebra.Etale.polynomial_eval_taylor_residual_pow}
\leanok
[Higher-order linear-Taylor residual at exponent $n+1$]
Let $g \in B[X]$, $c \in B$, and $\delta \in B$ with
$\delta \in ((\mathfrak m_A).\mathrm{map}(\mathrm{algebraMap}\,A\,B))^{n+1}$.
Then the Taylor residual at the base point $c$ in the direction $\delta$
satisfies
\[
g(c + \delta) - \bigl(g(c) + g'(c)\cdot\delta\bigr) \in
  ((\mathfrak m_A).\mathrm{map}(\mathrm{algebraMap}\,A\,B))^{2(n+1)}.
\]
\end{lemma}

\begin{proof}


\leanok
Direct from the single-variable identity
\texttt{Polynomial.binomExpansion}, which produces $q \in B$ with
$g(c + \delta) = g(c) + g'(c)\cdot\delta + q\cdot\delta^2$. Rearranging
gives the residual equal to $q\cdot\delta^2$. Since
$\delta \in I^{n+1}$ with $I := (\mathfrak m_A).\mathrm{map}
(\mathrm{algebraMap}\,A\,B)$, the membership
$\delta^2 \in (I^{n+1})^2 = I^{2(n+1)}$ follows from
\texttt{Ideal.pow\_mem\_pow}\,$\delta^{(n+1)}\,2$; multiplying by $q$
on the left (the ideal $I^{2(n+1)}$ is a $B$-submodule) closes the goal.
The base helper \texttt{basis\_expansion\_polynomial\_eval} (iter-068)
is the $n = 0$ instance of this lemma combined with a basis
decomposition for the direction $\delta = \sum_i \mathrm{algebraMap}
(\alpha_i)\,b_i$; the present lemma states the pure Taylor identity at
generic exponent.
\end{proof}

\section{Downstream consumers}

\Cref{thm:henselianRing-map-maximalIdeal} feeds the following
in-scope sorries (all currently typed sorries gated on this
instance). \emph{However}, because the corrected hypothesis
requires \texttt{Module.Finite A B}, the consumers in
\texttt{StrictlyHenselian.lean} and
\texttt{HenselianIdempotentLift.lean} — which currently work
with a general étale $B$ — will need a localisation step before
applying the instance. Specifically, for each maximal $\mathfrak
n$ of $B$ contracting to $\mathfrak m$, Stacks~04GG gives that
the localisation $B_{\mathfrak n}$ is finite étale over $A$;
applying \cref{thm:henselianRing-map-maximalIdeal} to
$B_{\mathfrak n}$ then gives the Henselian-pair property at
each fibre. Gluing the per-fibre lifts via the
orthogonal-idempotent decomposition closes the consumer.

\begin{itemize}
\item \texttt{Proetale/Mathlib/RingTheory/Etale/HenselianIdempotentLift.lean}
  (\texttt{exists\_completeOrthogonalIdempotents\_lift\_of\_henselian},
  L160): for each $i$, the naive lift $e_0 i$ lives in $B$; the
  Hensel-lifting of $X^2 - X$ at $e_0 i$ is performed in the
  finite-étale localisation $B_{\mathfrak n_i}$ at the
  corresponding maximal (where $B_{\mathfrak n_i}$ \emph{is}
  finite over $A$ by Stacks~04GG); the lifted idempotents in
  $B_{\mathfrak n_i}$ then patch back into idempotents in $B$
  via the residue-decomposition isomorphism. The derivative
  $2 e_0 i - 1$ is a unit mod $\mathfrak m B$ since under
  $\mathrm{eqv}$ it maps to $2 \cdot \pi_i - 1$ whose
  $i$-coordinate is $1$ and other coordinates are $-1$, all
  units.
\item \texttt{Proetale/Mathlib/RingTheory/Etale/StrictlyHenselian.lean}
  L682 (\texttt{exists\_idempotent\_lift\_isolating\_at\_maximal}'s
  horizontal-killing residual): the Stacks~04GH local-ness of each
  idempotent factor $B / (1 - e_{\mathrm{Lift}} i) B$ becomes
  available once the idempotents are true idempotents in $B$ —
  obtained by the same localise-then-glue pattern as above.
\item \texttt{StrictlyHenselian.lean} L1767 (\texttt{h\_polylift}
  inside \texttt{surjective\_standardEtalePair\_lift\_of\_etale}'s
  \S5 Gap A): the localisation $B_b = B[1/b]$ at the isolated
  idempotent is by construction a localisation at a single
  fibre, and so finite étale over $A$ at that fibre by
  Stacks~04GG; \cref{thm:henselianRing-map-maximalIdeal} then
  applies directly to $(B_b, \mathfrak m \cdot B_b)$.
\item \texttt{StrictlyHenselian.lean} L1814 (\S5 Gap B
  étale-Hensel-Nakayama close): same setting as L1767; the
  Hensel-lift in the finite-étale $B_b$ absorbs the $\mathfrak
  m B_b$-residual into a true algebra-image element of $B_b$.
\end{itemize}

The four consumer-wiring steps each split into (a) showing
\texttt{Module.Finite A B'} for the localised / fibre-restricted
$B'$, and (b) applying \cref{thm:henselianRing-map-maximalIdeal}.
Step (a) is the genuinely new piece introduced by the iter-053
counterexample finding; it requires the Mathlib infrastructure
\texttt{Algebra.Etale.localisation\_finite\_at\_maximal} or
equivalent (Stacks 04GG localised form). The iter-055+ planner
should scope this as a separate Mathlib helper before assigning
the wiring lanes.

\section{Implementation status}

\begin{itemize}
\item iter-052: opened the file
  \texttt{Proetale/Mathlib/RingTheory/Etale/HenselianPair.lean}
  with the headline instance signature in place (typed
  structural sorries).
\item iter-053: prover identified that the planner-authored
  signature is mathematically false (counterexample via
  \texttt{Algebra.Etale.of\_isLocalizationAway}); recorded the
  finding in \texttt{task\_results/}.
\item iter-054 plan: dispatched the \texttt{refactor} subagent
  (slug \texttt{amend-henselianpair-sig}) to add
  \texttt{[Module.Finite A B]} to the signature; the Lean file
  is now mathematically sound. This chapter restated to match.
\item iter-054 prover: closed \texttt{jac} sorry-free per the
  mechanical integral-going-up recipe above.
\item iter-055 prover (Acceptable-partial): \texttt{is\_henselian}
  decomposed sorry-free at the structural level — it delegates to
  a single named typed helper
  \texttt{exists\_root\_of\_eval\_mem\_of\_isUnit\_derivative\_quotient}
  whose signature matches the field verbatim. The Nakayama
  unit-upgrade glue (\texttt{maximalIdeal\_map\_le\_jacobson\_bot},
  \texttt{isUnit\_of\_isUnit\_quotient\_mk\_maximalIdeal\_map})
  landed sorry-free; the substantive root-finding step
  (Stages~1–3 above) is the single remaining typed sorry inside
  the named helper.
\item iter-056 plan: dispatched the \texttt{refactor} subagent
  (slug \texttt{add-noetherian-henselianpair}) to add
  \texttt{[IsNoetherianRing A]} to the headline instance and the
  three helpers, per the iter-055 prover finding that
  $\mathfrak m B$-adic separation (Stage~3 prerequisite) needs
  the Noetherian descent
  $\mathrm{IsNoetherianRing}\,A + \mathrm{Module.Finite}\,A\,B
  \Rightarrow \mathrm{IsNoetherianRing}\,B$ to invoke Mathlib's
  \texttt{Ideal.iInf\_pow\_smul\_eq\_bot\_of\_le\_jacobson}.
\item iter-056 prover (Acceptable-partial): closed
  $\mathfrak m B$-adic separation
  (\cref{lem:henselianPair-l1-separation}) sorry-free as
  \texttt{maximalIdeal\_map\_iInf\_pow\_eq\_bot} (10 LOC) and
  extracted Stage~2 Newton iteration as a typed named helper
  \texttt{exists\_seq\_lift\_of\_henselianPair} with body sorry.
\item iter-057 prover (Acceptable-full): closed
  \cref{lem:henselianPair-newton-step} sorry-free via
  $\Sigma'$-joint-state Newton recursion (\textasciitilde{}100
  LOC). The file is now down to exactly one sorry — the Stage~3
  convergence helper.
\item iter-058+ (current): closes the substantive Stage~3
  convergence helper via Stacks~04GE residue-product
  decomposition (orthogonal-idempotent lift in henselian local
  $A$ applied to the étale $B/\mathfrak m B \simeq \prod k_i$).
  Decomposes into sub-helpers L3a (idempotent lifting), L3b
  (product decomposition), L3c (per-factor Hensel-lift); plan
  acknowledges these are themselves new Mathlib-shaped
  infrastructure pieces and the prover may extract them as named
  typed sub-helpers. Once Stage~3 closes, iter-059+ wires the
  full \texttt{HenselianRing $B$ $\mathfrak m B$} instance into
  the four \S3-cluster consumers via the localise-then-glue
  pattern, together with the iter-054-identified consumer-side
  helper \texttt{Algebra.Etale.localisation\_finite\_at\_maximal}.
\end{itemize}
